---
title: "Producing JMP estimates"
subtitle: "WASH in healthcare facilities: Water, Sanitation and Hygiene indicators"
date: today
author:
  - name: Mohammad Shakkour
    orcid: 0009-0003-2050-7204
    email: shakkourm@who.int
    affiliation: 
      - name: World Health Organization Western Pacific Regional Office
        url: https://www.who.int/westernpacific
format:
  html:
    embed-resources: true
    toc: true
execute:
  warning: false
editor_options: 
  chunk_output_type: console
code-fold: true
license: "CC BY"
citation: 
  container-title: WASH in healthcare facilities
  url: https://github.com/Shakkour726
bibliography: references.bib
csl: apa.csl
editor: visual
---

# Introduction

This project aims to run an analysis to Water, Sanitation, Hygiene, Healthcare Waste Management and Environmental Cleaning (WASH) assessment in healthcare facilities. The analysis follows the World Health Organization (WHO) and the United Nations Children’s Fund (UNICEF) Joint Monitoring Programme for Water Supply, Sanitation and Hygiene (JMP) methodology to produce basic, limited and no service estimates. The sample is nationally representative according to the JMP methodology.

# Methods

The dataset contains the results of a WASH assessment in healthcare facilities carried out in one of the Member States. The WASH indicators are:

-   Water
-   Sanitation
-   Hygiene
-   Healthcare Waste Management
-   Environmental Cleaning

Calculating the estimates follows the [JMP methodology](https://washdata.org/reports/jmp-2017-methodology).

This project is running the analysis on water, sanitation and hygiene indicators **only**.

## Import and reading

Import libraries and reading the data

```{r}

#| label: load-packages
#| warning: false

# import libraries 
library(readr)
library(tidyverse)
library(dplyr)
library(ggthemes)
library(ggplot2)
library(knitr)

# reading file
data <- read_csv(here::here("data/processed/WinHCFs_microdata.csv"))
#glimpse(data)

```

## Stratifiers

Each stratifier includes more than 50 facilities. According to the JMP methodology, estimates can be produced for each stratifier.

```{r}

total_sample <- data |>
  count()

total_urban <- data |>
  filter(`urban/rural` == "Urban") |>
  count()

total_rural <- data |>
  filter(`urban/rural` == "Rural") |>
  count()

total_hospital <- data |>
  filter(`hospital/non-hospital` == "Hospital") |>
  count()

total_nonhospital <- data |>
  filter(`hospital/non-hospital` == "Non-Hospital") |>
  count()

total_public <- data |>
  filter(`public/private` == "Public") |>
  count()

total_private <- data |>
  filter(`public/private` == "Private") |>
  count()

```

## Water Data

### Improved sources

Calculating improved water source for each stratifier

```{r}

# calculating improved water source
GW1 <- data |>
  select(GW1, `urban/rural`, `hospital/non-hospital`, `public/private`)

total_gw1 <- GW1 |>
  count(GW1) |>
  mutate(Total = (n / sum(n))*100) |>
  select(GW1, Total)

urban_gw1 <- GW1 |>
  filter(`urban/rural`== "Urban") |>
  count(GW1) |>
  mutate(Urban = (n / sum(n))*100)|>
  select(GW1, Urban)

rural_gw1 <- GW1 |>
  filter(`urban/rural`== "Rural") |>
  count(GW1) |>
  mutate(Rural = (n / sum(n))*100)|>
  select(GW1, Rural)

hospital_gw1 <- GW1 |>
  filter(`hospital/non-hospital` == "Hospital") |>
  count(GW1) |>
  mutate(Hospital = (n / sum(n))*100)|>
  select(GW1, Hospital)

nonhospital_gw1 <- GW1 |>
  filter(`hospital/non-hospital` == "Non-Hospital") |>
  count(GW1) |>
  mutate(Nonhospital = (n / sum(n))*100)|>
  select(GW1, Nonhospital)

public_gw1 <- GW1 |>
  filter(`public/private` == "Public") |>
  count(GW1) |>
  mutate(Public = (n / sum(n))*100)|>
  select(GW1, Public)

private_gw1 <- GW1 |>
  filter(`public/private` == "Private") |>
  count(GW1) |>
  mutate(Private = (n / sum(n))*100)|>
  select(GW1, Private)

# joining
water_source <- total_gw1 |>
  left_join(urban_gw1) |>
  left_join(rural_gw1) |>
  left_join(hospital_gw1) |>
  left_join(nonhospital_gw1) |>
  left_join(public_gw1) |>
  left_join(private_gw1)

# replacing NA with zero
water_source <- water_source |>
  replace(is.na(water_source), 0)

# calculate improved
improved_water_source <- water_source |>
  filter(GW1 == "Piped Supply Outside the building" |
         GW1 == "Piped Supply inside the building" |
         GW1 == "Protected dug well" |
         GW1 == "Tanker truck" |
         GW1 == "Rainwater"    |
         GW1 == "Tube well/Borehole"
        ) |>
  summarise_at(vars(Total, Urban, Rural, Hospital, Nonhospital, Public, Private), sum)|>
  mutate(Indicator = "Improved water source") |>
  relocate(Indicator, .before = Total)

# calculate no service
no_service_water <- water_source |>
  filter(GW1 == "Don't Know" |
         GW1 == "No water source or patients bring water from home" |
         GW1 == "Other" |
         GW1 == "Unprotected dug well"
        ) |>
  summarise_at(vars(Total, Urban, Rural, Hospital, Nonhospital, Public, Private), sum) |>
  mutate(Indicator = "No service") |>
  relocate(Indicator, .before = Total)

```

### On premises

Calculating on premises indicator for each stratifier

```{r}

# calculate on premises
GW2 <- data |>
  select(GW2, `urban/rural`, `hospital/non-hospital`, `public/private`)

total_gw2 <- GW2 |>
  count(GW2) |>
  mutate(Total = (n / sum(n))*100) |>
  select(GW2, Total) |>
  filter(GW2 == "On premises")

urban_gw2 <- GW2 |>
  filter(`urban/rural`== "Urban") |>
  count(GW2) |>
  mutate(Urban = (n / sum(n))*100)|>
  select(GW2, Urban) |>
  filter(GW2 == "On premises")

rural_gw2 <- GW2 |>
  filter(`urban/rural`== "Rural") |>
  count(GW2) |>
  mutate(Rural = (n / sum(n))*100)|>
  select(GW2, Rural)|>
  filter(GW2 == "On premises")

hospital_gw2 <- GW2 |>
  filter(`hospital/non-hospital` == "Hospital") |>
  count(GW2) |>
  mutate(Hospital = (n / sum(n))*100)|>
  select(GW2, Hospital)|>
  filter(GW2 == "On premises")

nonhospital_gw2 <- GW2 |>
  filter(`hospital/non-hospital` == "Non-Hospital") |>
  count(GW2) |>
  mutate(Nonhospital = (n / sum(n))*100)|>
  select(GW2, Nonhospital)|>
  filter(GW2 == "On premises")

public_gw2 <- GW2 |>
  filter(`public/private` == "Public") |>
  count(GW2) |>
  mutate(Public = (n / sum(n))*100)|>
  select(GW2, Public)|>
  filter(GW2 == "On premises")

private_gw2 <- GW2 |>
  filter(`public/private` == "Private") |>
  count(GW2) |>
  mutate(Private = (n / sum(n))*100)|>
  select(GW2, Private)|>
  filter(GW2 == "On premises")

# joining
on_premises <- total_gw2 |>
  left_join(urban_gw2) |>
  left_join(rural_gw2) |>
  left_join(hospital_gw2) |>
  left_join(nonhospital_gw2) |>
  left_join(public_gw2) |>
  left_join(private_gw2) |>
  rename(Indicator = GW2)

```

### Availability

Calculating water availability indicator for each stratifier

```{r}

# calculate availability
GW3 <- data |>
  select(GW3, `urban/rural`, `hospital/non-hospital`, `public/private`)

total_gw3 <- GW3 |>
  count(GW3) |>
  mutate(Total = (n / sum(n))*100) |>
  select(GW3, Total) |>
  filter(GW3 == "Yes")

urban_gw3 <- GW3 |>
  filter(`urban/rural`== "Urban") |>
  count(GW3) |>
  mutate(Urban = (n / sum(n))*100)|>
  select(GW3, Urban) |>
  filter(GW3 == "Yes")

rural_gw3 <- GW3 |>
  filter(`urban/rural`== "Rural") |>
  count(GW3) |>
  mutate(Rural = (n / sum(n))*100)|>
  select(GW3, Rural)|>
  filter(GW3 == "Yes")

hospital_gw3 <- GW3 |>
  filter(`hospital/non-hospital` == "Hospital") |>
  count(GW3) |>
  mutate(Hospital = (n / sum(n))*100)|>
  select(GW3, Hospital)|>
  filter(GW3 == "Yes")

nonhospital_gw3 <- GW3 |>
  filter(`hospital/non-hospital` == "Non-Hospital") |>
  count(GW3) |>
  mutate(Nonhospital = (n / sum(n))*100)|>
  select(GW3, Nonhospital)|>
  filter(GW3 == "Yes")

public_gw3 <- GW3 |>
  filter(`public/private` == "Public") |>
  count(GW3) |>
  mutate(Public = (n / sum(n))*100)|>
  select(GW3, Public)|>
  filter(GW3 == "Yes")

private_gw3 <- GW3 |>
  filter(`public/private` == "Private") |>
  count(GW3) |>
  mutate(Private = (n / sum(n))*100)|>
  select(GW3, Private)|>
  filter(GW3 == "Yes")

# joining
available <- total_gw3 |>
  left_join(urban_gw3) |>
  left_join(rural_gw3) |>
  left_join(hospital_gw3) |>
  left_join(nonhospital_gw3) |>
  left_join(public_gw3) |>
  left_join(private_gw3) |>
  mutate(Indicator = "Available",
         GW3= NULL) |>
  relocate(Indicator, .before = Total)

```

### Improved and available

Calculating improved and available water indicator for each stratifier

```{r}

# calculating improved and available
GW13 <- data |>
  select(GW1, GW3, `urban/rural`, `hospital/non-hospital`, `public/private`)

total_gw13 <- GW13 |>
  filter(GW3 == "Yes") |>
  filter(GW1 == "Piped Supply Outside the building" |
         GW1 == "Piped Supply inside the building" |
         GW1 == "Protected dug well" |
         GW1 == "Tanker truck" |
         GW1 == "Rainwater"    |
         GW1 == "Tube well/Borehole") |>
  count(GW3) |>
  mutate(Total = (n / total_sample)*100) |>
  select(GW3, Total)

urban_gw13 <- GW13 |>
  filter(`urban/rural`== "Urban") |>
  filter(GW3 == "Yes") |>
  filter(GW1 == "Piped Supply Outside the building" |
         GW1 == "Piped Supply inside the building" |
         GW1 == "Protected dug well" |
         GW1 == "Tanker truck" |
         GW1 == "Rainwater"    |
         GW1 == "Tube well/Borehole") |>
  count(GW3) |>
  mutate(Urban = (n / total_urban)*100) |>
  select(GW3, Urban)

rural_gw13 <- GW13 |>
  filter(`urban/rural`== "Rural") |>
  filter(GW3 == "Yes") |>
  filter(GW1 == "Piped Supply Outside the building" |
         GW1 == "Piped Supply inside the building" |
         GW1 == "Protected dug well" |
         GW1 == "Tanker truck" |
         GW1 == "Rainwater"    |
         GW1 == "Tube well/Borehole") |>
  count(GW3) |>
  mutate(Rural = (n / total_rural)*100) |>
  select(GW3, Rural)


hospital_gw13 <- GW13 |>
  filter(`hospital/non-hospital`== "Hospital") |>
  filter(GW3 == "Yes") |>
  filter(GW1 == "Piped Supply Outside the building" |
         GW1 == "Piped Supply inside the building" |
         GW1 == "Protected dug well" |
         GW1 == "Tanker truck" |
         GW1 == "Rainwater"    |
         GW1 == "Tube well/Borehole") |>
  count(GW3) |>
  mutate(Hospital = (n / total_hospital)*100) |>
  select(GW3, Hospital)

nonhospital_gw13 <- GW13 |>
  filter(`hospital/non-hospital`== "Non-Hospital") |>
  filter(GW3 == "Yes") |>
  filter(GW1 == "Piped Supply Outside the building" |
         GW1 == "Piped Supply inside the building" |
         GW1 == "Protected dug well" |
         GW1 == "Tanker truck" |
         GW1 == "Rainwater"    |
         GW1 == "Tube well/Borehole") |>
  count(GW3) |>
  mutate(Nonhospital = (n / total_nonhospital)*100) |>
  select(GW3, Nonhospital)

public_gw13 <- GW13 |>
  filter(`public/private`== "Public") |>
  filter(GW3 == "Yes") |>
  filter(GW1 == "Piped Supply Outside the building" |
         GW1 == "Piped Supply inside the building" |
         GW1 == "Protected dug well" |
         GW1 == "Tanker truck" |
         GW1 == "Rainwater"    |
         GW1 == "Tube well/Borehole") |>
  count(GW3) |>
  mutate(Public = (n / total_public)*100) |>
  select(GW3, Public)

private_gw13 <- GW13 |>
  filter(`public/private`== "Private") |>
  filter(GW3 == "Yes") |>
  filter(GW1 == "Piped Supply Outside the building" |
         GW1 == "Piped Supply inside the building" |
         GW1 == "Protected dug well" |
         GW1 == "Tanker truck" |
         GW1 == "Rainwater"    |
         GW1 == "Tube well/Borehole") |>
  count(GW3) |>
  mutate(Private = (n / total_private)*100) |>
  select(GW3, Private)

# joining
improved_available <- total_gw13 |>
  left_join(urban_gw13) |>
  left_join(rural_gw13) |>
  left_join(hospital_gw13) |>
  left_join(nonhospital_gw13) |>
  left_join(public_gw13) |>
  left_join(private_gw13) |>
  mutate(Indicator = "Improved and available",
         GW3 = NULL,
         Total =  Total$n,
         Urban = Urban$n,
         Rural = Rural$n,
         Hospital = Hospital$n,
         Nonhospital = Nonhospital$n,
         Public = Public$n,
         Private = Private$n) |>
  relocate(Indicator, .before = Total)

```

### Improved and on premises

Calculating improved and on premises indicator for each stratifier

```{r}

#calculating improved and on premises
GW12 <- data |>
  select(GW1, GW2, `urban/rural`, `hospital/non-hospital`, `public/private`)

total_gw12 <- GW12 |>
  filter(GW2 == "On premises") |>
  filter(GW1 == "Piped Supply Outside the building" |
         GW1 == "Piped Supply inside the building" |
         GW1 == "Protected dug well" |
         GW1 == "Tanker truck" |
         GW1 == "Rainwater"    |
         GW1 == "Tube well/Borehole") |>
  count(GW2) |>
  mutate(Total = (n / total_sample)*100) |>
  select(GW2, Total)

urban_gw12 <- GW12 |>
  filter(`urban/rural`== "Urban") |>
  filter(GW2 == "On premises") |>
  filter(GW1 == "Piped Supply Outside the building" |
         GW1 == "Piped Supply inside the building" |
         GW1 == "Protected dug well" |
         GW1 == "Tanker truck" |
         GW1 == "Rainwater"    |
         GW1 == "Tube well/Borehole") |>
  count(GW2) |>
  mutate(Urban = (n / total_urban)*100) |>
  select(GW2, Urban)

rural_gw12 <- GW12 |>
  filter(`urban/rural`== "Rural") |>
  filter(GW2 == "On premises") |>
  filter(GW1 == "Piped Supply Outside the building" |
         GW1 == "Piped Supply inside the building" |
         GW1 == "Protected dug well" |
         GW1 == "Tanker truck" |
         GW1 == "Rainwater"    |
         GW1 == "Tube well/Borehole") |>
  count(GW2) |>
  mutate(Rural = (n / total_rural)*100) |>
  select(GW2, Rural)

hospital_gw12 <- GW12 |>
  filter(`hospital/non-hospital`== "Hospital") |>
  filter(GW2 == "On premises") |>
  filter(GW1 == "Piped Supply Outside the building" |
         GW1 == "Piped Supply inside the building" |
         GW1 == "Protected dug well" |
         GW1 == "Tanker truck" |
         GW1 == "Rainwater"    |
         GW1 == "Tube well/Borehole") |>
  count(GW2) |>
  mutate(Hospital = (n / total_hospital)*100) |>
  select(GW2, Hospital)

nonhospital_gw12 <- GW12 |>
  filter(`hospital/non-hospital`== "Non-Hospital") |>
  filter(GW2 == "On premises") |>
  filter(GW1 == "Piped Supply Outside the building" |
         GW1 == "Piped Supply inside the building" |
         GW1 == "Protected dug well" |
         GW1 == "Tanker truck" |
         GW1 == "Rainwater"    |
         GW1 == "Tube well/Borehole") |>
  count(GW2) |>
  mutate(Nonhospital = (n / total_nonhospital)*100) |>
  select(GW2, Nonhospital)

public_gw12 <- GW12 |>
  filter(`public/private`== "Public") |>
  filter(GW2 == "On premises") |>
  filter(GW1 == "Piped Supply Outside the building" |
         GW1 == "Piped Supply inside the building" |
         GW1 == "Protected dug well" |
         GW1 == "Tanker truck" |
         GW1 == "Rainwater"    |
         GW1 == "Tube well/Borehole") |>
  count(GW2) |>
  mutate(Public = (n / total_public)*100) |>
  select(GW2, Public)

private_gw12 <- GW12 |>
  filter(`public/private`== "Private") |>
  filter(GW2 == "On premises") |>
  filter(GW1 == "Piped Supply Outside the building" |
         GW1 == "Piped Supply inside the building" |
         GW1 == "Protected dug well" |
         GW1 == "Tanker truck" |
         GW1 == "Rainwater"    |
         GW1 == "Tube well/Borehole") |>
  count(GW2) |>
  mutate(Private = (n / total_private)*100) |>
  select(GW2, Private)

# joining
improved_onpremises <- total_gw12 |>
  left_join(urban_gw12) |>
  left_join(rural_gw12) |>
  left_join(hospital_gw12) |>
  left_join(nonhospital_gw12) |>
  left_join(public_gw12) |>
  left_join(private_gw12) |>
  mutate(Indicator = "Improved and on premises",
         GW2 = NULL,
         Total =  Total$n,
         Urban = Urban$n,
         Rural = Rural$n,
         Hospital = Hospital$n,
         Nonhospital = Nonhospital$n,
         Public = Public$n,
         Private = Private$n) |>
  relocate(Indicator, .before = Total)

```

### Basic and limited services

Calculating basic and limited service for each stratifier

```{r}

#calculating basic
GW123 <- data |>
  select(GW1, GW2, GW3, `urban/rural`, `hospital/non-hospital`, `public/private`)

basic_total <- GW123 |>
  filter(GW2 == "On premises") |>
  filter(GW3 == "Yes") |>
  filter(GW1 == "Piped Supply Outside the building" |
         GW1 == "Piped Supply inside the building" |
         GW1 == "Protected dug well" |
         GW1 == "Tanker truck" |
         GW1 == "Rainwater"    |
         GW1 == "Tube well/Borehole") |>
  count(GW2) |>
  mutate(Total = (n / total_sample)*100) |>
  select(GW2, Total)

basic_urban <- GW123 |>
  filter(`urban/rural`== "Urban") |>
  filter(GW2 == "On premises") |>
  filter(GW3 == "Yes") |>
  filter(GW1 == "Piped Supply Outside the building" |
         GW1 == "Piped Supply inside the building" |
         GW1 == "Protected dug well" |
         GW1 == "Tanker truck" |
         GW1 == "Rainwater"    |
         GW1 == "Tube well/Borehole") |>
  count(GW2) |>
  mutate(Urban = (n / total_urban)*100) |>
  select(GW2, Urban)

basic_rural <- GW123 |>
  filter(`urban/rural`== "Rural") |>
  filter(GW2 == "On premises") |>
  filter(GW3 == "Yes") |>
  filter(GW1 == "Piped Supply Outside the building" |
         GW1 == "Piped Supply inside the building" |
         GW1 == "Protected dug well" |
         GW1 == "Tanker truck" |
         GW1 == "Rainwater"    |
         GW1 == "Tube well/Borehole") |>
  count(GW2) |>
  mutate(Rural = (n / total_rural)*100) |>
  select(GW2, Rural)

basic_hospital <- GW123 |>
  filter(`hospital/non-hospital`== "Hospital") |>
  filter(GW2 == "On premises") |>
  filter(GW3 == "Yes") |>
  filter(GW1 == "Piped Supply Outside the building" |
         GW1 == "Piped Supply inside the building" |
         GW1 == "Protected dug well" |
         GW1 == "Tanker truck" |
         GW1 == "Rainwater"    |
         GW1 == "Tube well/Borehole") |>
  count(GW2) |>
  mutate(Hospital = (n / total_hospital)*100) |>
  select(GW2, Hospital)

basic_nonhospital <- GW123 |>
  filter(`hospital/non-hospital`== "Non-Hospital") |>
  filter(GW2 == "On premises") |>
  filter(GW3 == "Yes") |>
  filter(GW1 == "Piped Supply Outside the building" |
         GW1 == "Piped Supply inside the building" |
         GW1 == "Protected dug well" |
         GW1 == "Tanker truck" |
         GW1 == "Rainwater"    |
         GW1 == "Tube well/Borehole") |>
  count(GW2) |>
  mutate(Nonhospital = (n / total_nonhospital)*100) |>
  select(GW2, Nonhospital)

basic_public <- GW123 |>
  filter(`public/private`== "Public") |>
  filter(GW2 == "On premises") |>
  filter(GW3 == "Yes") |>
  filter(GW1 == "Piped Supply Outside the building" |
         GW1 == "Piped Supply inside the building" |
         GW1 == "Protected dug well" |
         GW1 == "Tanker truck" |
         GW1 == "Rainwater"    |
         GW1 == "Tube well/Borehole") |>
  count(GW2) |>
  mutate(Public = (n / total_public)*100) |>
  select(GW2, Public)

basic_private <- GW123 |>
  filter(`public/private`== "Private") |>
  filter(GW2 == "On premises") |>
  filter(GW3 == "Yes") |>
  filter(GW1 == "Piped Supply Outside the building" |
         GW1 == "Piped Supply inside the building" |
         GW1 == "Protected dug well" |
         GW1 == "Tanker truck" |
         GW1 == "Rainwater"    |
         GW1 == "Tube well/Borehole") |>
  count(GW2) |>
  mutate(Private = (n / total_private)*100) |>
  select(GW2, Private)

# joining
basic_water <- basic_total |>
  left_join(basic_urban) |>
  left_join(basic_rural) |>
  left_join(basic_hospital) |>
  left_join(basic_nonhospital) |>
  left_join(basic_public) |>
  left_join(basic_private) |>
  mutate(Indicator = "Basic",
         GW2 = NULL,
         Total =  Total$n,
         Urban = Urban$n,
         Rural = Rural$n,
         Hospital = Hospital$n,
         Nonhospital = Nonhospital$n,
         Public = Public$n,
         Private = Private$n) |>
  relocate(Indicator, .before = Total)

# calculating limited
limited_water <- tibble(
  Total = 100 - (basic_water[2] + no_service_water[2]),
  Urban = 100 - (basic_water[3] + no_service_water[3]),
  Rural = 100 - (basic_water[4] + no_service_water[4]),
  Hospital = 100 - (basic_water[5] + no_service_water[5]),
  Nonhospital = 100 - (basic_water[6] + no_service_water[6]),
  Public = 100 - (basic_water[7] + no_service_water[7]),
  Private = 100 - (basic_water[8] + no_service_water[8])) |>
  mutate(Indicator = "Limited",
         Total =  Total$Total,
         Urban = Urban$Urban,
         Rural = Rural$Rural,
         Hospital = Hospital$Hospital,
         Nonhospital = Nonhospital$Nonhospital,
         Public = Public$Public,
         Private = Private$Private) |>
  relocate(Indicator, .before = Total)

```

## Sanitation Data

### Improved facilities

Calculating improved sanitation facilities for each stratifier

```{r}

# calculating improved sanitation facility
GS1 <- data |>
  select(GS1, `urban/rural`, `hospital/non-hospital`, `public/private`)

total_gs1 <- GS1 |>
  count(GS1) |>
  mutate(Total = (n / sum(n))*100) |>
  select(GS1, Total)

urban_gs1 <- GS1 |>
  filter(`urban/rural`== "Urban") |>
  count(GS1) |>
  mutate(Urban = (n / sum(n))*100)|>
  select(GS1, Urban)

rural_gs1 <- GS1 |>
  filter(`urban/rural`== "Rural") |>
  count(GS1) |>
  mutate(Rural = (n / sum(n))*100)|>
  select(GS1, Rural)

hospital_gs1 <- GS1 |>
  filter(`hospital/non-hospital` == "Hospital") |>
  count(GS1) |>
  mutate(Hospital = (n / sum(n))*100)|>
  select(GS1, Hospital)

nonhospital_gs1 <- GS1 |>
  filter(`hospital/non-hospital` == "Non-Hospital") |>
  count(GS1) |>
  mutate(Nonhospital = (n / sum(n))*100)|>
  select(GS1, Nonhospital)

public_gs1 <- GS1 |>
  filter(`public/private` == "Public") |>
  count(GS1) |>
  mutate(Public = (n / sum(n))*100)|>
  select(GS1, Public)

private_gs1 <- GS1 |>
  filter(`public/private` == "Private") |>
  count(GS1) |>
  mutate(Private = (n / sum(n))*100)|>
  select(GS1, Private)

# joining
san_fac <- total_gs1 |>
  left_join(urban_gs1) |>
  left_join(rural_gs1) |>
  left_join(hospital_gs1) |>
  left_join(nonhospital_gs1) |>
  left_join(public_gs1) |>
  left_join(private_gs1)

# replacing NA with zero
san_fac <- san_fac |>
  replace(is.na(san_fac), 0)

# calculate improved
improved_san_fac <- san_fac |>
  filter(GS1 == "Flush/Pour-flush toilet to tank or pit" |
         GS1 == "Pit latrine with slab" |
         GS1 == "pit latrine with slab") |>
  summarise_at(vars(Total, Urban, Rural, Hospital, Nonhospital, Public, Private), sum)|>
  mutate(Indicator = "Improved sanitation facility") |>
  relocate(Indicator, .before = Total)

# calculate no service
no_service_san <- san_fac |>
  filter(GS1 == "Pit latrine without slab" |
         GS1 == "other") |>
  summarise_at(vars(Total, Urban, Rural, Hospital, Nonhospital, Public, Private), sum) |>
  mutate(Indicator = "No service") |>
  relocate(Indicator, .before = Total)

```

### Usable facilities

Calculating usable sanitation facilities for each stratifier

```{r}

# calculate usable facilities
GS2 <- data |>
  select(GS2, `urban/rural`, `hospital/non-hospital`, `public/private`)

total_gs2 <- GS2 |>
  count(GS2) |>
  mutate(Total = (n / sum(n))*100) |>
  select(GS2, Total) |>
  filter(GS2 == "Yes")

urban_gs2 <- GS2 |>
  filter(`urban/rural`== "Urban") |>
  count(GS2) |>
  mutate(Urban = (n / sum(n))*100)|>
  select(GS2, Urban) |>
  filter(GS2 == "Yes")

rural_gs2 <- GS2 |>
  filter(`urban/rural`== "Rural") |>
  count(GS2) |>
  mutate(Rural = (n / sum(n))*100)|>
  select(GS2, Rural)|>
  filter(GS2 == "Yes")

hospital_gs2 <- GS2 |>
  filter(`hospital/non-hospital` == "Hospital") |>
  count(GS2) |>
  mutate(Hospital = (n / sum(n))*100)|>
  select(GS2, Hospital)|>
  filter(GS2 == "Yes")

nonhospital_gs2 <- GS2 |>
  filter(`hospital/non-hospital` == "Non-Hospital") |>
  count(GS2) |>
  mutate(Nonhospital = (n / sum(n))*100)|>
  select(GS2, Nonhospital)|>
  filter(GS2 == "Yes")

public_gs2 <- GS2 |>
  filter(`public/private` == "Public") |>
  count(GS2) |>
  mutate(Public = (n / sum(n))*100)|>
  select(GS2, Public)|>
  filter(GS2 == "Yes")

private_gs2 <- GS2 |>
  filter(`public/private` == "Private") |>
  count(GS2) |>
  mutate(Private = (n / sum(n))*100)|>
  select(GS2, Private)|>
  filter(GS2 == "Yes")

# joining
san_usable <- total_gs2 |>
  left_join(urban_gs2) |>
  left_join(rural_gs2) |>
  left_join(hospital_gs2) |>
  left_join(nonhospital_gs2) |>
  left_join(public_gs2) |>
  left_join(private_gs2) |>
  mutate(Indicator = "Usable sanitation facility",
         GS2 = NULL) |>
  relocate(Indicator, .before = Total)

```

### Facilities dedicated to staff

Calculating facilities dedicated for staff for each stratifier

```{r}

# calculate facilities dedicated for staff
GS3 <- data |>
  select(GS3, `urban/rural`, `hospital/non-hospital`, `public/private`)

total_gs3 <- GS3 |>
  count(GS3) |>
  mutate(Total = (n / sum(n))*100) |>
  select(GS3, Total) |>
  filter(GS3 == "Yes")

urban_gs3 <- GS3 |>
  filter(`urban/rural`== "Urban") |>
  count(GS3) |>
  mutate(Urban = (n / sum(n))*100)|>
  select(GS3, Urban) |>
  filter(GS3 == "Yes")

rural_gs3 <- GS3 |>
  filter(`urban/rural`== "Rural") |>
  count(GS3) |>
  mutate(Rural = (n / sum(n))*100)|>
  select(GS3, Rural)|>
  filter(GS3 == "Yes")

hospital_gs3 <- GS3 |>
  filter(`hospital/non-hospital` == "Hospital") |>
  count(GS3) |>
  mutate(Hospital = (n / sum(n))*100)|>
  select(GS3, Hospital)|>
  filter(GS3 == "Yes")

nonhospital_gs3 <- GS3 |>
  filter(`hospital/non-hospital` == "Non-Hospital") |>
  count(GS3) |>
  mutate(Nonhospital = (n / sum(n))*100)|>
  select(GS3, Nonhospital)|>
  filter(GS3 == "Yes")

public_gs3 <- GS3 |>
  filter(`public/private` == "Public") |>
  count(GS3) |>
  mutate(Public = (n / sum(n))*100)|>
  select(GS3, Public)|>
  filter(GS3 == "Yes")

private_gs3 <- GS3 |>
  filter(`public/private` == "Private") |>
  count(GS3) |>
  mutate(Private = (n / sum(n))*100)|>
  select(GS3, Private)|>
  filter(GS3 == "Yes")

# joining
san_staff <- total_gs3 |>
  left_join(urban_gs3) |>
  left_join(rural_gs3) |>
  left_join(hospital_gs3) |>
  left_join(nonhospital_gs3) |>
  left_join(public_gs3) |>
  left_join(private_gs3) |>
  mutate(Indicator = "Dedicated to staff",
         GS3 = NULL) |>
  relocate(Indicator, .before = Total)

```

### Sex-separated facilities

Calculating sex-separated facilities for each stratifier

```{r}

# calculate sex-separated facilities
GS4 <- data |>
  select(GS4, `urban/rural`, `hospital/non-hospital`, `public/private`)

total_gs4 <- GS4 |>
  count(GS4) |>
  mutate(Total = (n / sum(n))*100) |>
  select(GS4, Total) |>
  filter(GS4 == "Yes")

urban_gs4 <- GS4 |>
  filter(`urban/rural`== "Urban") |>
  count(GS4) |>
  mutate(Urban = (n / sum(n))*100)|>
  select(GS4, Urban) |>
  filter(GS4 == "Yes")

rural_gs4 <- GS4 |>
  filter(`urban/rural`== "Rural") |>
  count(GS4) |>
  mutate(Rural = (n / sum(n))*100)|>
  select(GS4, Rural)|>
  filter(GS4 == "Yes")

hospital_gs4 <- GS4 |>
  filter(`hospital/non-hospital` == "Hospital") |>
  count(GS4) |>
  mutate(Hospital = (n / sum(n))*100)|>
  select(GS4, Hospital)|>
  filter(GS4 == "Yes")

nonhospital_gs4 <- GS4 |>
  filter(`hospital/non-hospital` == "Non-Hospital") |>
  count(GS4) |>
  mutate(Nonhospital = (n / sum(n))*100)|>
  select(GS4, Nonhospital)|>
  filter(GS4 == "Yes")

public_gs4 <- GS4 |>
  filter(`public/private` == "Public") |>
  count(GS4) |>
  mutate(Public = (n / sum(n))*100)|>
  select(GS4, Public)|>
  filter(GS4 == "Yes")

private_gs4 <- GS4 |>
  filter(`public/private` == "Private") |>
  count(GS4) |>
  mutate(Private = (n / sum(n))*100)|>
  select(GS4, Private)|>
  filter(GS4 == "Yes")

# joining
san_sex <- total_gs4 |>
  left_join(urban_gs4) |>
  left_join(rural_gs4) |>
  left_join(hospital_gs4) |>
  left_join(nonhospital_gs4) |>
  left_join(public_gs4) |>
  left_join(private_gs4) |>
  mutate(Indicator = "Sex-separated facility",
         GS4 = NULL) |>
  relocate(Indicator, .before = Total)

```

### Menstrual hygiene facilities

Calculating menstrual hygiene facilities for each stratifier

```{r}

# calculate menstrual hygiene facilities
GS5 <- data |>
  select(GS5, `urban/rural`, `hospital/non-hospital`, `public/private`)

total_gs5 <- GS5 |>
  count(GS5) |>
  mutate(Total = (n / sum(n))*100) |>
  select(GS5, Total) |>
  filter(GS5 == "Yes")

urban_gs5 <- GS5 |>
  filter(`urban/rural`== "Urban") |>
  count(GS5) |>
  mutate(Urban = (n / sum(n))*100)|>
  select(GS5, Urban) |>
  filter(GS5 == "Yes")

rural_gs5 <- GS5 |>
  filter(`urban/rural`== "Rural") |>
  count(GS5) |>
  mutate(Rural = (n / sum(n))*100)|>
  select(GS5, Rural)|>
  filter(GS5 == "Yes")

hospital_gs5 <- GS5 |>
  filter(`hospital/non-hospital` == "Hospital") |>
  count(GS5) |>
  mutate(Hospital = (n / sum(n))*100)|>
  select(GS5, Hospital)|>
  filter(GS5 == "Yes")

nonhospital_gs5 <- GS5 |>
  filter(`hospital/non-hospital` == "Non-Hospital") |>
  count(GS5) |>
  mutate(Nonhospital = (n / sum(n))*100)|>
  select(GS5, Nonhospital)|>
  filter(GS5 == "Yes")

public_gs5 <- GS5 |>
  filter(`public/private` == "Public") |>
  count(GS5) |>
  mutate(Public = (n / sum(n))*100)|>
  select(GS5, Public)|>
  filter(GS5 == "Yes")

private_gs5 <- GS5 |>
  filter(`public/private` == "Private") |>
  count(GS5) |>
  mutate(Private = (n / sum(n))*100)|>
  select(GS5, Private)|>
  filter(GS5 == "Yes")

# joining
san_mhm <- total_gs5 |>
  left_join(urban_gs5) |>
  left_join(rural_gs5) |>
  left_join(hospital_gs5) |>
  left_join(nonhospital_gs5) |>
  left_join(public_gs5) |>
  left_join(private_gs5) |>
  mutate(Indicator = "Menstrual hygiene faciliy",
         GS5 = NULL) |>
  relocate(Indicator, .before = Total)

```

### Accessible for limited mobility

Calculating facilities accessible for limited mobility for each stratifier

```{r}

# calculate facilities accessible for limited mobility
GS6 <- data |>
  select(GS6, `urban/rural`, `hospital/non-hospital`, `public/private`)

total_gs6 <- GS6 |>
  count(GS6) |>
  mutate(Total = (n / sum(n))*100) |>
  select(GS6, Total) |>
  filter(GS6 == "yes")

urban_gs6 <- GS6 |>
  filter(`urban/rural`== "Urban") |>
  count(GS6) |>
  mutate(Urban = (n / sum(n))*100)|>
  select(GS6, Urban) |>
  filter(GS6 == "yes")

rural_gs6 <- GS6 |>
  filter(`urban/rural`== "Rural") |>
  count(GS6) |>
  mutate(Rural = (n / sum(n))*100)|>
  select(GS6, Rural)|>
  filter(GS6 == "yes")

hospital_gs6 <- GS6 |>
  filter(`hospital/non-hospital` == "Hospital") |>
  count(GS6) |>
  mutate(Hospital = (n / sum(n))*100)|>
  select(GS6, Hospital)|>
  filter(GS6 == "yes")

nonhospital_gs6 <- GS6 |>
  filter(`hospital/non-hospital` == "Non-Hospital") |>
  count(GS6) |>
  mutate(Nonhospital = (n / sum(n))*100)|>
  select(GS6, Nonhospital)|>
  filter(GS6 == "yes")

public_gs6 <- GS6 |>
  filter(`public/private` == "Public") |>
  count(GS6) |>
  mutate(Public = (n / sum(n))*100)|>
  select(GS6, Public)|>
  filter(GS6 == "yes")

private_gs6 <- GS6 |>
  filter(`public/private` == "Private") |>
  count(GS6) |>
  mutate(Private = (n / sum(n))*100)|>
  select(GS6, Private)|>
  filter(GS6 == "yes")

# joining
san_mobility <- total_gs6 |>
  left_join(urban_gs6) |>
  left_join(rural_gs6) |>
  left_join(hospital_gs6) |>
  left_join(nonhospital_gs6) |>
  left_join(public_gs6) |>
  left_join(private_gs6) |>
  mutate(Indicator = "Accessible for limited mobility",
         GS6 = NULL) |>
  relocate(Indicator, .before = Total)

```

### Improved and usable

Calculating improved and usable indicator for each stratifier

```{r}

#calculating improved and usable
GS12 <- data |>
  select(GS1, GS2, `urban/rural`, `hospital/non-hospital`, `public/private`)

total_gs12 <- GS12 |>
  filter(GS2 == "Yes") |>
  filter(GS1 == "Flush/Pour-flush toilet to tank or pit" |
         GS1 == "Pit latrine with slab" |
         GS1 == "pit latrine with slab") |>
  count(GS2) |>
  mutate(Total = (n / total_sample)*100) |>
  select(GS2, Total)

urban_gs12 <- GS12 |>
  filter(`urban/rural`== "Urban") |>
  filter(GS2 == "Yes") |>
  filter(GS1 == "Flush/Pour-flush toilet to tank or pit" |
         GS1 == "Pit latrine with slab" |
         GS1 == "pit latrine with slab") |>
  count(GS2) |>
  mutate(Urban = (n / total_urban)*100) |>
  select(GS2, Urban)

rural_gs12 <- GS12 |>
  filter(`urban/rural`== "Rural") |>
  filter(GS2 == "Yes") |>
  filter(GS1 == "Flush/Pour-flush toilet to tank or pit" |
         GS1 == "Pit latrine with slab" |
         GS1 == "pit latrine with slab") |>
  count(GS2) |>
  mutate(Rural = (n / total_rural)*100) |>
  select(GS2, Rural)

hospital_gs12 <- GS12 |>
  filter(`hospital/non-hospital`== "Hospital") |>
  filter(GS2 == "Yes") |>
  filter(GS1 == "Flush/Pour-flush toilet to tank or pit" |
         GS1 == "Pit latrine with slab" |
         GS1 == "pit latrine with slab") |>
  count(GS2) |>
  mutate(Hospital = (n / total_hospital)*100) |>
  select(GS2, Hospital)

nonhospital_gs12 <- GS12 |>
  filter(`hospital/non-hospital`== "Non-Hospital") |>
  filter(GS2 == "Yes") |>
  filter(GS1 == "Flush/Pour-flush toilet to tank or pit" |
         GS1 == "Pit latrine with slab" |
         GS1 == "pit latrine with slab") |>
  count(GS2) |>
  mutate(Nonhospital = (n / total_nonhospital)*100) |>
  select(GS2, Nonhospital)

public_gs12 <- GS12 |>
  filter(`public/private`== "Public") |>
  filter(GS2 == "Yes") |>
  filter(GS1 == "Flush/Pour-flush toilet to tank or pit" |
         GS1 == "Pit latrine with slab" |
         GS1 == "pit latrine with slab") |>
  count(GS2) |>
  mutate(Public = (n / total_public)*100) |>
  select(GS2, Public)

private_gs12 <- GS12 |>
  filter(`public/private`== "Private") |>
  filter(GS2 == "Yes") |>
  filter(GS1 == "Flush/Pour-flush toilet to tank or pit" |
         GS1 == "Pit latrine with slab" |
         GS1 == "pit latrine with slab") |>
  count(GS2) |>
  mutate(Private = (n / total_private)*100) |>
  select(GS2, Private)

# joining
improved_usable <- total_gs12 |>
  left_join(urban_gs12) |>
  left_join(rural_gs12) |>
  left_join(hospital_gs12) |>
  left_join(nonhospital_gs12) |>
  left_join(public_gs12) |>
  left_join(private_gs12) |>
  mutate(Indicator = "Improved and usable",
         GS2 = NULL,
         Total =  Total$n,
         Urban = Urban$n,
         Rural = Rural$n,
         Hospital = Hospital$n,
         Nonhospital = Nonhospital$n,
         Public = Public$n,
         Private = Private$n) |>
  relocate(Indicator, .before = Total)

```

### Improved and staff

Calculating improved and staff indicator for each stratifier

```{r}

#calculating improved and staff
GS13 <- data |>
  select(GS1, GS3, `urban/rural`, `hospital/non-hospital`, `public/private`)

total_gs13 <- GS13 |>
  filter(GS3 == "Yes") |>
  filter(GS1 == "Flush/Pour-flush toilet to tank or pit" |
         GS1 == "Pit latrine with slab" |
         GS1 == "pit latrine with slab") |>
  count(GS3) |>
  mutate(Total = (n / total_sample)*100) |>
  select(GS3, Total)

urban_gs13 <- GS13 |>
  filter(`urban/rural`== "Urban") |>
  filter(GS3 == "Yes") |>
  filter(GS1 == "Flush/Pour-flush toilet to tank or pit" |
         GS1 == "Pit latrine with slab" |
         GS1 == "pit latrine with slab") |>
  count(GS3) |>
  mutate(Urban = (n / total_urban)*100) |>
  select(GS3, Urban)

rural_gs13 <- GS13 |>
  filter(`urban/rural`== "Rural") |>
  filter(GS3 == "Yes") |>
  filter(GS1 == "Flush/Pour-flush toilet to tank or pit" |
         GS1 == "Pit latrine with slab" |
         GS1 == "pit latrine with slab") |>
  count(GS3) |>
  mutate(Rural = (n / total_rural)*100) |>
  select(GS3, Rural)

hospital_gs13 <- GS13 |>
  filter(`hospital/non-hospital`== "Hospital") |>
  filter(GS3 == "Yes") |>
  filter(GS1 == "Flush/Pour-flush toilet to tank or pit" |
         GS1 == "Pit latrine with slab" |
         GS1 == "pit latrine with slab") |>
  count(GS3) |>
  mutate(Hospital = (n / total_hospital)*100) |>
  select(GS3, Hospital)

nonhospital_gs13 <- GS13 |>
  filter(`hospital/non-hospital`== "Non-Hospital") |>
  filter(GS3 == "Yes") |>
  filter(GS1 == "Flush/Pour-flush toilet to tank or pit" |
         GS1 == "Pit latrine with slab" |
         GS1 == "pit latrine with slab") |>
  count(GS3) |>
  mutate(Nonhospital = (n / total_nonhospital)*100) |>
  select(GS3, Nonhospital)

public_gs13 <- GS13 |>
  filter(`public/private`== "Public") |>
  filter(GS3 == "Yes") |>
  filter(GS1 == "Flush/Pour-flush toilet to tank or pit" |
         GS1 == "Pit latrine with slab" |
         GS1 == "pit latrine with slab") |>
  count(GS3) |>
  mutate(Public = (n / total_public)*100) |>
  select(GS3, Public)

private_gs13 <- GS13 |>
  filter(`public/private`== "Private") |>
  filter(GS3 == "Yes") |>
  filter(GS1 == "Flush/Pour-flush toilet to tank or pit" |
         GS1 == "Pit latrine with slab" |
         GS1 == "pit latrine with slab") |>
  count(GS3) |>
  mutate(Private = (n / total_private)*100) |>
  select(GS3, Private)

# joining
improved_staff <- total_gs13 |>
  left_join(urban_gs13) |>
  left_join(rural_gs13) |>
  left_join(hospital_gs13) |>
  left_join(nonhospital_gs13) |>
  left_join(public_gs13) |>
  left_join(private_gs13) |>
  mutate(Indicator = "Improved and staff",
         GS3 = NULL,
         Total =  Total$n,
         Urban = Urban$n,
         Rural = Rural$n,
         Hospital = Hospital$n,
         Nonhospital = Nonhospital$n,
         Public = Public$n,
         Private = Private$n) |>
  relocate(Indicator, .before = Total)

```

### Improved and sex-separated

Calculating improved and sex-separated indicator for each stratifier

```{r}

#calculating improved and sex-separated
GS14 <- data |>
  select(GS1, GS4, `urban/rural`, `hospital/non-hospital`, `public/private`)

total_gs14 <- GS14 |>
  filter(GS4 == "Yes") |>
  filter(GS1 == "Flush/Pour-flush toilet to tank or pit" |
         GS1 == "Pit latrine with slab" |
         GS1 == "pit latrine with slab") |>
  count(GS4) |>
  mutate(Total = (n / total_sample)*100) |>
  select(GS4, Total)

urban_gs14 <- GS14 |>
  filter(`urban/rural`== "Urban") |>
  filter(GS4 == "Yes") |>
  filter(GS1 == "Flush/Pour-flush toilet to tank or pit" |
         GS1 == "Pit latrine with slab" |
         GS1 == "pit latrine with slab") |>
  count(GS4) |>
  mutate(Urban = (n / total_urban)*100) |>
  select(GS4, Urban)

rural_gs14 <- GS14 |>
  filter(`urban/rural`== "Rural") |>
  filter(GS4 == "Yes") |>
  filter(GS1 == "Flush/Pour-flush toilet to tank or pit" |
         GS1 == "Pit latrine with slab" |
         GS1 == "pit latrine with slab") |>
  count(GS4) |>
  mutate(Rural = (n / total_rural)*100) |>
  select(GS4, Rural)

hospital_gs14 <- GS14 |>
  filter(`hospital/non-hospital`== "Hospital") |>
  filter(GS4 == "Yes") |>
  filter(GS1 == "Flush/Pour-flush toilet to tank or pit" |
         GS1 == "Pit latrine with slab" |
         GS1 == "pit latrine with slab") |>
  count(GS4) |>
  mutate(Hospital = (n / total_hospital)*100) |>
  select(GS4, Hospital)

nonhospital_gs14 <- GS14 |>
  filter(`hospital/non-hospital`== "Non-Hospital") |>
  filter(GS4 == "Yes") |>
  filter(GS1 == "Flush/Pour-flush toilet to tank or pit" |
         GS1 == "Pit latrine with slab" |
         GS1 == "pit latrine with slab") |>
  count(GS4) |>
  mutate(Nonhospital = (n / total_nonhospital)*100) |>
  select(GS4, Nonhospital)

public_gs14 <- GS14 |>
  filter(`public/private`== "Public") |>
  filter(GS4 == "Yes") |>
  filter(GS1 == "Flush/Pour-flush toilet to tank or pit" |
         GS1 == "Pit latrine with slab" |
         GS1 == "pit latrine with slab") |>
  count(GS4) |>
  mutate(Public = (n / total_public)*100) |>
  select(GS4, Public)

private_gs14 <- GS14 |>
  filter(`public/private`== "Private") |>
  filter(GS4 == "Yes") |>
  filter(GS1 == "Flush/Pour-flush toilet to tank or pit" |
         GS1 == "Pit latrine with slab" |
         GS1 == "pit latrine with slab") |>
  count(GS4) |>
  mutate(Private = (n / total_private)*100) |>
  select(GS4, Private)

# joining
improved_sex <- total_gs14 |>
  left_join(urban_gs14) |>
  left_join(rural_gs14) |>
  left_join(hospital_gs14) |>
  left_join(nonhospital_gs14) |>
  left_join(public_gs14) |>
  left_join(private_gs14) |>
  mutate(Indicator = "Improved and sex-separated",
         GS4 = NULL,
         Total =  Total$n,
         Urban = Urban$n,
         Rural = Rural$n,
         Hospital = Hospital$n,
         Nonhospital = Nonhospital$n,
         Public = Public$n,
         Private = Private$n) |>
  relocate(Indicator, .before = Total)

```

### Improved and menstrual hygiene

Calculating improved and menstrual hygiene indicator for each stratifier

```{r}

#calculating improved and menstrual hygiene
GS15 <- data |>
  select(GS1, GS5, `urban/rural`, `hospital/non-hospital`, `public/private`)

total_gs15 <- GS15 |>
  filter(GS5 == "Yes") |>
  filter(GS1 == "Flush/Pour-flush toilet to tank or pit" |
         GS1 == "Pit latrine with slab" |
         GS1 == "pit latrine with slab") |>
  count(GS5) |>
  mutate(Total = (n / total_sample)*100) |>
  select(GS5, Total)

urban_gs15 <- GS15 |>
  filter(`urban/rural`== "Urban") |>
  filter(GS5 == "Yes") |>
  filter(GS1 == "Flush/Pour-flush toilet to tank or pit" |
         GS1 == "Pit latrine with slab" |
         GS1 == "pit latrine with slab") |>
  count(GS5) |>
  mutate(Urban = (n / total_urban)*100) |>
  select(GS5, Urban)

rural_gs15 <- GS15 |>
  filter(`urban/rural`== "Rural") |>
  filter(GS5 == "Yes") |>
  filter(GS1 == "Flush/Pour-flush toilet to tank or pit" |
         GS1 == "Pit latrine with slab" |
         GS1 == "pit latrine with slab") |>
  count(GS5) |>
  mutate(Rural = (n / total_rural)*100) |>
  select(GS5, Rural)

hospital_gs15 <- GS15 |>
  filter(`hospital/non-hospital`== "Hospital") |>
  filter(GS5 == "Yes") |>
  filter(GS1 == "Flush/Pour-flush toilet to tank or pit" |
         GS1 == "Pit latrine with slab" |
         GS1 == "pit latrine with slab") |>
  count(GS5) |>
  mutate(Hospital = (n / total_hospital)*100) |>
  select(GS5, Hospital)

nonhospital_gs15 <- GS15 |>
  filter(`hospital/non-hospital`== "Non-Hospital") |>
  filter(GS5 == "Yes") |>
  filter(GS1 == "Flush/Pour-flush toilet to tank or pit" |
         GS1 == "Pit latrine with slab" |
         GS1 == "pit latrine with slab") |>
  count(GS5) |>
  mutate(Nonhospital = (n / total_nonhospital)*100) |>
  select(GS5, Nonhospital)

public_gs15 <- GS15 |>
  filter(`public/private`== "Public") |>
  filter(GS5 == "Yes") |>
  filter(GS1 == "Flush/Pour-flush toilet to tank or pit" |
         GS1 == "Pit latrine with slab" |
         GS1 == "pit latrine with slab") |>
  count(GS5) |>
  mutate(Public = (n / total_public)*100) |>
  select(GS5, Public)

private_gs15 <- GS15 |>
  filter(`public/private`== "Private") |>
  filter(GS5 == "Yes") |>
  filter(GS1 == "Flush/Pour-flush toilet to tank or pit" |
         GS1 == "Pit latrine with slab" |
         GS1 == "pit latrine with slab") |>
  count(GS5) |>
  mutate(Private = (n / total_private)*100) |>
  select(GS5, Private)

# joining
improved_mhm <- total_gs15 |>
  left_join(urban_gs15) |>
  left_join(rural_gs15) |>
  left_join(hospital_gs15) |>
  left_join(nonhospital_gs15) |>
  left_join(public_gs15) |>
  left_join(private_gs15) |>
  mutate(Indicator = "Improved and menstrual hygiene",
         GS5 = NULL,
         Total =  Total$n,
         Urban = Urban$n,
         Rural = Rural$n,
         Hospital = Hospital$n,
         Nonhospital = Nonhospital$n,
         Public = Public$n,
         Private = Private$n) |>
  relocate(Indicator, .before = Total)

```

### Improved and limited mobility

Calculating improved and limited mobility indicator for each stratifier

```{r}

#calculating improved and limited mobility
GS16 <- data |>
  select(GS1, GS6, `urban/rural`, `hospital/non-hospital`, `public/private`)

total_gs16 <- GS16 |>
  filter(GS6 == "yes") |>
  filter(GS1 == "Flush/Pour-flush toilet to tank or pit" |
         GS1 == "Pit latrine with slab" |
         GS1 == "pit latrine with slab") |>
  count(GS6) |>
  mutate(Total = (n / total_sample)*100) |>
  select(GS6, Total)

urban_gs16 <- GS16 |>
  filter(`urban/rural`== "Urban") |>
  filter(GS6 == "yes") |>
  filter(GS1 == "Flush/Pour-flush toilet to tank or pit" |
         GS1 == "Pit latrine with slab" |
         GS1 == "pit latrine with slab") |>
  count(GS6) |>
  mutate(Urban = (n / total_urban)*100) |>
  select(GS6, Urban)

rural_gs16 <- GS16 |>
  filter(`urban/rural`== "Rural") |>
  filter(GS6 == "yes") |>
  filter(GS1 == "Flush/Pour-flush toilet to tank or pit" |
         GS1 == "Pit latrine with slab" |
         GS1 == "pit latrine with slab") |>
  count(GS6) |>
  mutate(Rural = (n / total_rural)*100) |>
  select(GS6, Rural)

hospital_gs16 <- GS16 |>
  filter(`hospital/non-hospital`== "Hospital") |>
  filter(GS6 == "yes") |>
  filter(GS1 == "Flush/Pour-flush toilet to tank or pit" |
         GS1 == "Pit latrine with slab" |
         GS1 == "pit latrine with slab") |>
  count(GS6) |>
  mutate(Hospital = (n / total_hospital)*100) |>
  select(GS6, Hospital)

nonhospital_gs16 <- GS16 |>
  filter(`hospital/non-hospital`== "Non-Hospital") |>
  filter(GS6 == "yes") |>
  filter(GS1 == "Flush/Pour-flush toilet to tank or pit" |
         GS1 == "Pit latrine with slab" |
         GS1 == "pit latrine with slab") |>
  count(GS6) |>
  mutate(Nonhospital = (n / total_nonhospital)*100) |>
  select(GS6, Nonhospital)

public_gs16 <- GS16 |>
  filter(`public/private`== "Public") |>
  filter(GS6 == "yes") |>
  filter(GS1 == "Flush/Pour-flush toilet to tank or pit" |
         GS1 == "Pit latrine with slab" |
         GS1 == "pit latrine with slab") |>
  count(GS6) |>
  mutate(Public = (n / total_public)*100) |>
  select(GS6, Public)

private_gs16 <- GS16 |>
  filter(`public/private`== "Private") |>
  filter(GS6 == "yes") |>
  filter(GS1 == "Flush/Pour-flush toilet to tank or pit" |
         GS1 == "Pit latrine with slab" |
         GS1 == "pit latrine with slab") |>
  count(GS6) |>
  mutate(Private = (n / total_private)*100) |>
  select(GS6, Private)

# joining
improved_mobility <- total_gs16 |>
  left_join(urban_gs16) |>
  left_join(rural_gs16) |>
  left_join(hospital_gs16) |>
  left_join(nonhospital_gs16) |>
  left_join(public_gs16) |>
  left_join(private_gs16) |>
  mutate(Indicator = "Improved and limited mobility",
         GS6 = NULL,
         Total =  Total$n,
         Urban = Urban$n,
         Rural = Rural$n,
         Hospital = Hospital$n,
         Nonhospital = Nonhospital$n,
         Public = Public$n,
         Private = Private$n) |>
  relocate(Indicator, .before = Total)

```

### Basic and limited services

Calculating basic and limited service for each stratifier

```{r}

#calculating basic
GS <- data |>
  select(GS1, GS2, GS3, GS4, GS5, GS6, `urban/rural`, `hospital/non-hospital`, `public/private`)

basic_total <- GS |>
  filter(GS2 == "Yes") |>
  filter(GS3 == "Yes") |>
  filter(GS4 == "Yes") |>
  filter(GS5 == "Yes") |>
  filter(GS6 == "yes") |>
  filter(GS1 == "Flush/Pour-flush toilet to tank or pit" |
         GS1 == "Pit latrine with slab" |
         GS1 == "pit latrine with slab") |>
  count(GS2) |>
  mutate(Total = (n / total_sample)*100) |>
  select(GS2, Total)

basic_urban <- GS |>
  filter(`urban/rural`== "Urban") |>
  filter(GS2 == "Yes") |>
  filter(GS3 == "Yes") |>
  filter(GS4 == "Yes") |>
  filter(GS5 == "Yes") |>
  filter(GS6 == "yes") |>
  filter(GS1 == "Flush/Pour-flush toilet to tank or pit" |
         GS1 == "Pit latrine with slab" |
         GS1 == "pit latrine with slab") |>
  count(GS2) |>
  mutate(Urban = (n / total_urban)*100) |>
  select(GS2, Urban)

basic_rural <- GS |>
  filter(`urban/rural`== "Rural") |>
  filter(GS2 == "Yes") |>
  filter(GS3 == "Yes") |>
  filter(GS4 == "Yes") |>
  filter(GS5 == "Yes") |>
  filter(GS6 == "yes") |>
  filter(GS1 == "Flush/Pour-flush toilet to tank or pit" |
         GS1 == "Pit latrine with slab" |
         GS1 == "pit latrine with slab") |>
  count(GS2) |>
  mutate(Rural = (n / total_rural)*100) |>
  select(GS2, Rural)

basic_hospital <- GS |>
  filter(`hospital/non-hospital`== "Hospital") |>
  filter(GS2 == "Yes") |>
  filter(GS3 == "Yes") |>
  filter(GS4 == "Yes") |>
  filter(GS5 == "Yes") |>
  filter(GS6 == "yes") |>
  filter(GS1 == "Flush/Pour-flush toilet to tank or pit" |
         GS1 == "Pit latrine with slab" |
         GS1 == "pit latrine with slab") |>
  count(GS2) |>
  mutate(Hospital = (n / total_hospital)*100) |>
  select(GS2, Hospital)

basic_nonhospital <- GS |>
  filter(`hospital/non-hospital`== "Non-Hospital") |>
  filter(GS2 == "Yes") |>
  filter(GS3 == "Yes") |>
  filter(GS4 == "Yes") |>
  filter(GS5 == "Yes") |>
  filter(GS6 == "yes") |>
  filter(GS1 == "Flush/Pour-flush toilet to tank or pit" |
         GS1 == "Pit latrine with slab" |
         GS1 == "pit latrine with slab") |>
  count(GS2) |>
  mutate(Nonhospital = (n / total_nonhospital)*100) |>
  select(GS2, Nonhospital)

basic_public <- GS |>
  filter(`public/private`== "Public") |>
  filter(GS2 == "Yes") |>
  filter(GS3 == "Yes") |>
  filter(GS4 == "Yes") |>
  filter(GS5 == "Yes") |>
  filter(GS6 == "yes") |>
  filter(GS1 == "Flush/Pour-flush toilet to tank or pit" |
         GS1 == "Pit latrine with slab" |
         GS1 == "pit latrine with slab") |>
  count(GS2) |>
  mutate(Public = (n / total_public)*100) |>
  select(GS2, Public)

basic_private <- GS |>
  filter(`public/private`== "Private") |>
  filter(GS2 == "Yes") |>
  filter(GS3 == "Yes") |>
  filter(GS4 == "Yes") |>
  filter(GS5 == "Yes") |>
  filter(GS6 == "yes") |>
  filter(GS1 == "Flush/Pour-flush toilet to tank or pit" |
         GS1 == "Pit latrine with slab" |
         GS1 == "pit latrine with slab") |>
  count(GS2) |>
  mutate(Private = (n / total_private)*100) |>
  select(GS2, Private)

# joining
basic_san <- basic_total |>
  left_join(basic_urban) |>
  left_join(basic_rural) |>
  left_join(basic_hospital) |>
  left_join(basic_nonhospital) |>
  left_join(basic_public) |>
  left_join(basic_private) |>
  mutate(Indicator = "Basic",
         GS2 = NULL,
         Total =  Total$n,
         Urban = Urban$n,
         Rural = Rural$n,
         Hospital = Hospital$n,
         Nonhospital = Nonhospital$n,
         Public = Public$n,
         Private = Private$n) |>
  relocate(Indicator, .before = Total)

# calculating limited
limited_san <- tibble(
  Total = 100 - (basic_san[2] + no_service_san[2]),
  Urban = 100 - (basic_san[3] + no_service_san[3]),
  Rural = 100 - (basic_san[4] + no_service_san[4]),
  Hospital = 100 - (basic_san[5] + no_service_san[5]),
  Nonhospital = 100 - (basic_san[6] + no_service_san[6]),
  Public = 100 - (basic_san[7] + no_service_san[7]),
  Private = 100 - (basic_san[8] + no_service_san[8])) |>
  mutate(Indicator = "Limited",
         Total =  Total$Total,
         Urban = Urban$Urban,
         Rural = Rural$Rural,
         Hospital = Hospital$Hospital,
         Nonhospital = Nonhospital$Nonhospital,
         Public = Public$Public,
         Private = Private$Private) |>
  relocate(Indicator, .before = Total)

```

## Hygiene Data

### Points of care

Calculating handwashing facilities with water and soap at points of care for each stratifier

```{r}

# calculating handwashing facilities with water and soap at points of care
GH1 <- data |>
  select(GH1, `urban/rural`, `hospital/non-hospital`, `public/private`) |>
  filter(GH1 == "Yes")

total_gh1 <- GH1 |>
  count(GH1) |>
  mutate(Total = (n / total_sample)*100) |>
  select(GH1, Total)

urban_gh1 <- GH1 |>
  filter(`urban/rural`== "Urban") |>
  count(GH1) |>
  mutate(Urban = (n / total_urban)*100)|>
  select(GH1, Urban)

rural_gh1 <- GH1 |>
  filter(`urban/rural`== "Rural") |>
  count(GH1) |>
  mutate(Rural = (n / total_rural)*100)|>
  select(GH1, Rural)

hospital_gh1 <- GH1 |>
  filter(`hospital/non-hospital` == "Hospital") |>
  count(GH1) |>
  mutate(Hospital = (n / total_hospital)*100)|>
  select(GH1, Hospital)

nonhospital_gh1 <- GH1 |>
  filter(`hospital/non-hospital` == "Non-Hospital") |>
  count(GH1) |>
  mutate(Nonhospital = (n / total_nonhospital)*100)|>
  select(GH1, Nonhospital)

public_gh1 <- GH1 |>
  filter(`public/private` == "Public") |>
  count(GH1) |>
  mutate(Public = (n / total_public)*100)|>
  select(GH1, Public)

private_gh1 <- GH1 |>
  filter(`public/private` == "Private") |>
  count(GH1) |>
  mutate(Private = (n / total_private)*100)|>
  select(GH1, Private)

# joining and replace NA with zero 
pointsofcare <- total_gh1 |>
  left_join(urban_gh1) |>
  left_join(rural_gh1) |>
  left_join(hospital_gh1) |>
  left_join(nonhospital_gh1) |>
  left_join(public_gh1) |>
  left_join(private_gh1) |>
  mutate(Indicator = "Handwashing facilities with water and soap at points of care",
         GH1 = NULL,
         Total =  Total$n,
         Urban = Urban$n,
         Rural = Rural$n,
         Hospital = Hospital$n,
         Nonhospital = Nonhospital$n,
         Public = Public$n,
         Private = Private$n) |>
  relocate(Indicator, .before = Total)

```

### Toilets

Calculating handwashing facilities within 5 meters of toilets for each stratifier

```{r}

# calculating handwashing facilities within 5 metres of toilets
GH2 <- data |>
  select(GH2, `urban/rural`, `hospital/non-hospital`, `public/private`)

total_gh2 <- GH2 |>
  filter(GH2 == "Yes" |
         GH2 == "No, there are handwashing facilities near the toilets but lacking soap and /or water") |>
  count(GH2) |>
  summarise_at(vars(n), sum) |>
  mutate(GH2 = "Yes",
         Total = (n / total_sample) * 100,
         n = NULL)

urban_gh2 <- GH2 |>
  filter(`urban/rural`== "Urban",
         GH2 == "Yes" |
         GH2 == "No, there are handwashing facilities near the toilets but lacking soap and /or water") |>
  count(GH2) |>
  summarise_at(vars(n), sum) |>
  mutate(GH2 = "Yes",
         Urban = (n / total_urban) * 100,
         n = NULL)

rural_gh2 <- GH2 |>
  filter(`urban/rural`== "Rural",
         GH2 == "Yes" |
         GH2 == "No, there are handwashing facilities near the toilets but lacking soap and /or water") |>
  count(GH2) |>
  summarise_at(vars(n), sum) |>
  mutate(GH2 = "Yes",
         Rural = (n / total_rural) * 100,
         n = NULL)

hospital_gh2 <- GH2 |>
  filter(`hospital/non-hospital` == "Hospital",
         GH2 == "Yes" |
         GH2 == "No, there are handwashing facilities near the toilets but lacking soap and /or water") |>
  count(GH2) |>
  summarise_at(vars(n), sum) |>
  mutate(GH2 = "Yes",
         Hospital = (n / total_hospital) * 100,
         n = NULL)

nonhospital_gh2 <- GH2 |>
  filter(`hospital/non-hospital` == "Non-Hospital",
         GH2 == "Yes" |
         GH2 == "No, there are handwashing facilities near the toilets but lacking soap and /or water") |>
  count(GH2) |>
  summarise_at(vars(n), sum) |>
  mutate(GH2 = "Yes",
         Nonhospital = (n / total_nonhospital) * 100,
         n = NULL)

public_gh2 <- GH2 |>
  filter(`public/private` == "Public",
         GH2 == "Yes" |
         GH2 == "No, there are handwashing facilities near the toilets but lacking soap and /or water") |>
  count(GH2) |>
  summarise_at(vars(n), sum) |>
  mutate(GH2 = "Yes",
         Public = (n / total_public) * 100,
         n = NULL)

private_gh2 <- GH2 |>
  filter(`public/private` == "Private",
         GH2 == "Yes" |
         GH2 == "No, there are handwashing facilities near the toilets but lacking soap and /or water") |>
  count(GH2) |>
  summarise_at(vars(n), sum) |>
  mutate(GH2 = "Yes",
         Private = (n / total_private) * 100,
         n = NULL)

# joining
toilets <- total_gh2 |>
  left_join(urban_gh2) |>
  left_join(rural_gh2) |>
  left_join(hospital_gh2) |>
  left_join(nonhospital_gh2) |>
  left_join(public_gh2) |>
  left_join(private_gh2) |>
  mutate(Indicator = "Handwashing facilities within 5 metres of toilets",
         GH2 = NULL,
         Total =  Total$n,
         Urban = Urban$n,
         Rural = Rural$n,
         Hospital = Hospital$n,
         Nonhospital = Nonhospital$n,
         Public = Public$n,
         Private = Private$n) |>
  relocate(Indicator, .before = Total)

# calculating handwashing facilities within 5 metres of toilets with water and soap
GH2_fun <- data |>
  select(GH2, `urban/rural`, `hospital/non-hospital`, `public/private`) |>
  filter(GH2 == "Yes")

total_gh2_fun <- GH2_fun |>
  count(GH2) |>
  mutate(Total = (n / total_sample)*100) |>
  select(GH2, Total)

urban_gh2_fun <- GH2_fun |>
  filter(`urban/rural`== "Urban") |>
  count(GH2) |>
  mutate(Urban = (n / total_urban)*100)|>
  select(GH2, Urban)

rural_gh2_fun <- GH2_fun |>
  filter(`urban/rural`== "Rural") |>
  count(GH2) |>
  mutate(Rural = (n / total_rural)*100)|>
  select(GH2, Rural)

hospital_gh2_fun <- GH2_fun |>
  filter(`hospital/non-hospital` == "Hospital") |>
  count(GH2) |>
  mutate(Hospital = (n / total_hospital)*100)|>
  select(GH2, Hospital)

nonhospital_gh2_fun <- GH2_fun |>
  filter(`hospital/non-hospital` == "Non-Hospital") |>
  count(GH2) |>
  mutate(Nonhospital = (n / total_nonhospital)*100)|>
  select(GH2, Nonhospital)

public_gh2_fun <- GH2_fun |>
  filter(`public/private` == "Public") |>
  count(GH2) |>
  mutate(Public = (n / total_public)*100)|>
  select(GH2, Public)

private_gh2_fun <- GH2_fun |>
  filter(`public/private` == "Private") |>
  count(GH2) |>
  mutate(Private = (n / total_private)*100)|>
  select(GH2, Private)

# joining
toilets_fun <- total_gh2_fun |>
  left_join(urban_gh2_fun) |>
  left_join(rural_gh2_fun) |>
  left_join(hospital_gh2_fun) |>
  left_join(nonhospital_gh2_fun) |>
  left_join(public_gh2_fun) |>
  left_join(private_gh2_fun) |>
  mutate(Indicator = "Handwashing facilities within 5 metres of toilets with water and soap",
         GH2 = NULL,
         Total =  Total$n,
         Urban = Urban$n,
         Rural = Rural$n,
         Hospital = Hospital$n,
         Nonhospital = Nonhospital$n,
         Public = Public$n,
         Private = Private$n) |>
  relocate(Indicator, .before = Total)

```

### Basic and limited services

Calculating basic and limited service for each stratifier

```{r}

# calculating basic
GH12 <- data |>
  select(GH1, GH2, `urban/rural`, `hospital/non-hospital`, `public/private`) |>
  filter(GH1 == "Yes" &
         GH2 == "Yes")

total_gh12 <- GH12 |>
  count(GH2) |>
  mutate(Total = (n / total_sample)*100) |>
  select(GH2, Total)

urban_gh12 <- GH12 |>
  filter(`urban/rural`== "Urban") |>
  count(GH2) |>
  mutate(Urban = (n / total_urban)*100)|>
  select(GH2, Urban)

rural_gh12 <- GH12 |>
  filter(`urban/rural`== "Rural") |>
  count(GH2) |>
  mutate(Rural = (n / total_rural)*100)|>
  select(GH2, Rural)

hospital_gh12 <- GH12 |>
  filter(`hospital/non-hospital` == "Hospital") |>
  count(GH2) |>
  mutate(Hospital = (n / total_hospital)*100)|>
  select(GH2, Hospital)

nonhospital_gh12 <- GH12 |>
  filter(`hospital/non-hospital` == "Non-Hospital") |>
  count(GH2) |>
  mutate(Nonhospital = (n / total_nonhospital)*100)|>
  select(GH2, Nonhospital)

public_gh12 <- GH12 |>
  filter(`public/private` == "Public") |>
  count(GH2) |>
  mutate(Public = (n / total_public)*100)|>
  select(GH2, Public)

private_gh12 <- GH12 |>
  filter(`public/private` == "Private") |>
  count(GH2) |>
  mutate(Private = (n / total_private)*100)|>
  select(GH2, Private)

# joining
basic_hygiene <- total_gh12 |>
  left_join(urban_gh12) |>
  left_join(rural_gh12) |>
  left_join(hospital_gh12) |>
  left_join(nonhospital_gh12) |>
  left_join(public_gh12) |>
  left_join(private_gh12) |>
  mutate(Indicator = "Basic",
         GH2 = NULL,
         Total =  Total$n,
         Urban = Urban$n,
         Rural = Rural$n,
         Hospital = Hospital$n,
         Nonhospital = Nonhospital$n,
         Public = Public$n,
         Private = Private$n) |>
  relocate(Indicator, .before = Total)

# calculating no serivce
GH12_no <- data |>
  select(GH1, GH2, `urban/rural`, `hospital/non-hospital`, `public/private`)

total_gh12_no <- GH12_no |>
  filter(GH1 != "Yes" &
         GH2 != "Yes") |>
  count(GH2) |>
  summarise_at(vars(n), sum) |>
  mutate(GH2 = "Yes",
         Total = (n / total_sample) * 100,
         n = NULL)

urban_gh12_no <- GH12_no |>
  filter(`urban/rural`== "Urban") |>
  filter(GH1 != "Yes" &
         GH2 != "Yes") |>
  count(GH2) |>
  summarise_at(vars(n), sum) |>
  mutate(GH2 = "Yes",
         Urban = (n / total_urban) * 100,
         n = NULL)

rural_gh12_no <- GH12_no |>
  filter(`urban/rural`== "Rural") |>
  filter(GH1 != "Yes" &
         GH2 != "Yes") |>
  count(GH2) |>
  summarise_at(vars(n), sum) |>
  mutate(GH2 = "Yes",
         Rural = (n / total_rural) * 100,
         n = NULL)

hospital_gh12_no <- GH12_no |>
  filter(`hospital/non-hospital` == "Hospital") |>
  filter(GH1 != "Yes" &
         GH2 != "Yes") |>
  count(GH2) |>
  summarise_at(vars(n), sum) |>
  mutate(GH2 = "Yes",
         Hospital = (n / total_hospital) * 100,
         n = NULL)

nonhospital_gh12_no <- GH12_no |>
  filter(`hospital/non-hospital` == "Non-Hospital") |>
  filter(GH1 != "Yes" &
         GH2 != "Yes") |>
  count(GH2) |>
  summarise_at(vars(n), sum) |>
  mutate(GH2 = "Yes",
         Nonhospital = (n / total_nonhospital) * 100,
         n = NULL)

public_gh12_no <- GH12_no |>
  filter(`public/private` == "Public") |>
  filter(GH1 != "Yes" &
         GH2 != "Yes") |>
  count(GH2) |>
  summarise_at(vars(n), sum) |>
  mutate(GH2 = "Yes",
         Public = (n / total_public) * 100,
         n = NULL)

private_gh12_no <- GH12_no |>
  filter(`public/private` == "Private") |>
  filter(GH1 != "Yes" &
         GH2 != "Yes") |>
  count(GH2) |>
  summarise_at(vars(n), sum) |>
  mutate(GH2 = "Yes",
         Private = (n / total_private) * 100,
         n = NULL)

# joining
no_service_hygiene <- total_gh12_no |>
  left_join(urban_gh12_no) |>
  left_join(rural_gh12_no) |>
  left_join(hospital_gh12_no) |>
  left_join(nonhospital_gh12_no) |>
  left_join(public_gh12_no) |>
  left_join(private_gh12_no) |>
  mutate(Indicator = "No service",
         GH2 = NULL,
         Total =  Total$n,
         Urban = Urban$n,
         Rural = Rural$n,
         Hospital = Hospital$n,
         Nonhospital = Nonhospital$n,
         Public = Public$n,
         Private = Private$n) |>
  relocate(Indicator, .before = Total)

# calculating limited
limited_hygiene <- tibble(
  Total = 100 - (basic_hygiene[2] + no_service_hygiene[2]),
  Urban = 100 - (basic_hygiene[3] + no_service_hygiene[3]),
  Rural = 100 - (basic_hygiene[4] + no_service_hygiene[4]),
  Hospital = 100 - (basic_hygiene[5] + no_service_hygiene[5]),
  Nonhospital = 100 - (basic_hygiene[6] + no_service_hygiene[6]),
  Public = 100 - (basic_hygiene[7] + no_service_hygiene[7]),
  Private = 100 - (basic_hygiene[8] + no_service_hygiene[8])) |>
  mutate(Indicator = "Limited",
         Total =  Total$Total,
         Urban = Urban$Urban,
         Rural = Rural$Rural,
         Hospital = Hospital$Hospital,
         Nonhospital = Nonhospital$Nonhospital,
         Public = Public$Public,
         Private = Private$Private) |>
  relocate(Indicator, .before = Total)

```

# Results

## Water indicators

See @tbl-water summarizes all water indicators resulted from the analysis.

Significant disparities in water indicators exist in terms of the types of facilities; facilities located in urban areas as well as hospitals are more likely to meet the criteria for basic water service than facilities located in rural areas and smaller health care facilities, respectively.

```{r}
#| label: tbl-water
#| tbl-cap: Water indicators

improved_water_source |>
  full_join(on_premises) |>
  full_join(available) |>
  full_join(improved_available) |>
  full_join(improved_onpremises) |>
  full_join(basic_water) |>
  full_join(limited_water) |>
  full_join(no_service_water) |>
  kable(digits = 1)

```

## Sanitation indicators

See @tbl-sanitation summarizes all sanitation indicators resulted from the analysis.

Significant disparities in sanitation indicators exist in terms of the types of facilities; facilities located in urban areas as well as hospitals are more likely to meet the criteria for basic sanitation service than facilities located in rural areas and smaller health care facilities, respectively.

```{r}
#| label: tbl-sanitation
#| tbl-cap: Sanitation indicators

improved_san_fac |>
  full_join(san_usable) |>
  full_join(san_staff) |>
  full_join(san_sex) |>
  full_join(san_mhm) |>
  full_join(san_mobility) |>
  full_join(improved_usable) |>
  full_join(improved_staff) |>
  full_join(improved_sex) |>
  full_join(improved_mhm) |>
  full_join(improved_mobility) |>
  full_join(basic_san) |>
  full_join(limited_san) |>
  full_join(no_service_san) |>
  kable(digits = 1)

```

## Hygiene indicators

See @tbl-hygiene summarizes all hygiene indicators resulted from the analysis.

Significant disparities in hygiene indicators exist in terms of the types of facilities; facilities located in urban areas as well as hospitals are more likely to meet the criteria for basic hygiene service than facilities located in rural areas and smaller health care facilities, respectively.

```{r}
#| label: tbl-hygiene
#| tbl-cap: Hygiene indicators

pointsofcare |>
  full_join(toilets) |>
  full_join(toilets_fun) |>
  full_join(basic_hygiene) |>
  full_join(limited_hygiene) |>
  full_join(no_service_hygiene) |>
  kable(digits = 1)

```

# Conclusion

## Basic water ladder

See @fig-water illustrates basic water ladder resulted from the analysis. Significant disparities are depicted in the basic water ladders across all settings.

```{r}
#| label: fig-water
#| fig-cap: Basic water ladder

# create object
basic_water_ladder <-
  (basic_water) |>
  full_join(limited_water) |>
  full_join(no_service_water)

# pivot
basic_water_ladder <-
  basic_water_ladder |>
  pivot_longer(cols = Total:Private,
             names_to = "Setting",
             values_to = "Percent")

# reorder indicators
basic_water_ladder$Indicator <- factor(basic_water_ladder$Indicator, levels = c("No service", "Limited", "Basic"))

# reorder settings
basic_water_ladder$Setting <- factor(basic_water_ladder$Setting, levels = c("Total", "Urban", "Rural", "Hospital", "Nonhospital", "Public", "Private"))

# Assigning white for basic in geom_text
basic_water_ladder <- basic_water_ladder |>
  mutate(Color = case_when(
    Indicator == "Basic" ~ "white",
    .default = "black"))

# plot water ladders
ggplot(data = basic_water_ladder,
       mapping = aes(x = Setting,
                     y = Percent,
                     fill = Indicator)) +
  geom_col() +
  geom_text(aes(label = round(Percent, 0)),
            position = position_stack(vjust = 0.5),
            size = 3,
            color = basic_water_ladder$Color) +
  theme_minimal() +
  labs(title = "Water Indicator",
       subtitle = "JMP basic water service ladder",
       x = "Setting",
       y = "Proportion of healthcare facilities (%)") +
  scale_fill_manual(
    values = c("No service" = "#FEBC11", "Limited" = "#FFF176", "Basic" = "#00B8EC"),
    breaks = c("No service", "Limited","Basic")) +
  guides(fill = guide_legend(title = NULL)) +
  theme(panel.grid.minor.y = element_blank(),
        panel.grid.major.x = element_blank())

```

## Basic sanitation ladder

See @fig-sanitation illustrates basic sanitation ladder resulted from the analysis. Basic estimates are too low across all settings.

```{r}
#| label: fig-sanitation
#| fig-cap: Basic sanitation ladder

#create object
basic_san_ladder <-
  (basic_san) |>
  full_join(limited_san) |>
  full_join(no_service_san)

#pivot
basic_san_ladder <-
  basic_san_ladder |>
  pivot_longer(cols = Total:Private,
             names_to = "Setting",
             values_to = "Percent")

#reorder indicators
basic_san_ladder$Indicator <- factor(basic_san_ladder$Indicator, levels = c("No service", "Limited", "Basic"))

#reorder settings
basic_san_ladder$Setting <- factor(basic_san_ladder$Setting, levels = c("Total", "Urban", "Rural", "Hospital", "Nonhospital", "Public", "Private"))

# Assigning white for basic in geom_text
basic_san_ladder <- basic_san_ladder |>
  mutate(Color = case_when(
    Indicator == "Basic" ~ "white",
    .default = "black"))

# plot sanitation ladders
ggplot(data = basic_san_ladder,
       mapping = aes(x = Setting,
                     y = Percent,
                     fill = Indicator)) +
  geom_col() +
  geom_text(aes(label = round(Percent, 0)),
            position = position_stack(vjust = 0.5),
            size = 3,
            color = basic_san_ladder$Color) +
  theme_minimal() +
  labs(title = "Sanitation Indicator",
       subtitle = "JMP basic sanitation service ladder",
       x = "Setting",
       y = "Proportion of healthcare facilities (%)") +
  scale_fill_manual(
    values = c("No service" = "#FEBC11", "Limited" = "#FFF176", "Basic" = "#66BB6A"),
    breaks = c("No service", "Limited","Basic")) +
  guides(fill = guide_legend(title = NULL)) +
  theme(panel.grid.minor.y = element_blank(),
        panel.grid.major.x = element_blank())

```

## Basic hygiene ladder

See @fig-hygiene illustrates basic hygiene ladder resulted from the analysis. Significant disparities are depicted in the basic hygiene ladders across all settings.

```{r}
#| label: fig-hygiene
#| fig-cap: Basic hygiene ladder

# create object
basic_hygiene_ladder <-
  (basic_hygiene) |>
  full_join(limited_hygiene) |>
  full_join(no_service_hygiene)

# pivot
basic_hygiene_ladder <-
  basic_hygiene_ladder |>
  pivot_longer(cols = Total:Private,
             names_to = "Setting",
             values_to = "Percent")

# reorder indicators
basic_hygiene_ladder$Indicator <- factor(basic_hygiene_ladder$Indicator, levels = c("No service", "Limited", "Basic"))

# reorder settings
basic_hygiene_ladder$Setting <- factor(basic_hygiene_ladder$Setting, levels = c("Total", "Urban", "Rural", "Hospital", "Nonhospital", "Public", "Private"))

# Assigning white for basic in geom_text
basic_hygiene_ladder <- basic_hygiene_ladder |>
  mutate(Color = case_when(
    Indicator == "Basic" ~ "white",
    .default = "black"))

# plot hygiene ladders
ggplot(data = basic_hygiene_ladder,
       mapping = aes(x = Setting,
                     y = Percent,
                     fill = Indicator)) +
  geom_col() +
  geom_text(aes(label = round(Percent, 0)),
            position = position_stack(vjust = 0.5),
            size = 3,
            color = basic_hygiene_ladder$Color) +
  theme_minimal() +
  labs(title = "Hygiene Indicator",
       subtitle = "JMP basic hygiene service ladder",
       x = "Setting",
       y = "Proportion of healthcare facilities (%)") +
  scale_fill_manual(
    values = c("No service" = "#FEBC11", "Limited" = "#FFF176", "Basic" = "#AB47BC"),
    breaks = c("No service", "Limited","Basic")) +
  guides(fill = guide_legend(title = NULL)) +
  theme(panel.grid.minor.y = element_blank(),
        panel.grid.major.x = element_blank())
```

# References

1.  @WHOUNICEF2017
2.  @WHOUNICEF2019
