---
title: "Producing JMP estimates"
subtitle: "WASH in healthcare facilities: Water and Hygiene indicators"
date: today
author:
  - name: Mohammad Shakkour
    orcid: 0009-0003-2050-7204
    email: shakkourm@who.int
    affiliation: 
      - name: World Health Organization Western Pacific Regional Office
        url: https://www.who.int/westernpacific
format:
  html:
    embed-resources: true
    toc: true
execute:
  warning: false
editor_options: 
  chunk_output_type: console
code-fold: true
license: "CC BY"
citation: 
  container-title: WASH in healthcare facilities
  url: https://github.com/Shakkour726
bibliography: references.bib
csl: apa.csl
editor: visual
---

# Introduction

This project aims to run an analysis to Water, Sanitation, Hygiene, Healthcare Waste Management and Environmental Cleaning (WASH) assessment in healthcare facilities. The analysis follows the World Health Organization (WHO) and the United Nations Children’s Fund (UNICEF) Joint Monitoring Programme for Water Supply, Sanitation and Hygiene (JMP) methodology to produce basic, limited and no service estimates. The sample is nationally representative according to the JMP methodology.

# Methods

The dataset contains the results of a WASH assessment in healthcare facilities carried out in one of the Member States. The WASH indicators are:

-   Water
-   Sanitation
-   Hygiene
-   Healthcare Waste Management
-   Environmental Cleaning

Calculating the estimates follows the [JMP methodology](https://washdata.org/reports/jmp-2017-methodology).

Phase 1 of the project will run the analysis on water and hygiene indicators **only**.

## Import and reading

Import libraries and reading the data

```{r}

#| label: load-packages
#| warning: false

# import libraries 
library(readr)
library(tidyverse)
library(dplyr)
library(ggthemes)
library(ggplot2)
library(knitr)

# reading file
data <- read_csv(here::here("data/processed/WinHCFs_microdata.csv"))
#glimpse(data)

```

## Stratifiers

Each stratifier includes more than 50 facilities. According to the JMP methodology, estimates can be produced for each stratifier.

```{r}

total_sample <- data |>
  count()

total_urban <- data |>
  filter(`urban/rural` == "Urban") |>
  count()

total_rural <- data |>
  filter(`urban/rural` == "Rural") |>
  count()

total_hospital <- data |>
  filter(`hospital/non-hospital` == "Hospital") |>
  count()

total_nonhospital <- data |>
  filter(`hospital/non-hospital` == "Non-Hospital") |>
  count()

total_public <- data |>
  filter(`public/private` == "Public") |>
  count()

total_private <- data |>
  filter(`public/private` == "Private") |>
  count()

```

## Water Data

### Improved sources

Calculating improved water source for each stratifier

```{r}

# calculating improved water source
GW1 <- data |>
  select(GW1, `urban/rural`, `hospital/non-hospital`, `public/private`)

total_gw1 <- GW1 |>
  count(GW1) |>
  mutate(Total = (n / sum(n))*100) |>
  select(GW1, Total)

urban_gw1 <- GW1 |>
  filter(`urban/rural`== "Urban") |>
  count(GW1) |>
  mutate(Urban = (n / sum(n))*100)|>
  select(GW1, Urban)

rural_gw1 <- GW1 |>
  filter(`urban/rural`== "Rural") |>
  count(GW1) |>
  mutate(Rural = (n / sum(n))*100)|>
  select(GW1, Rural)

hospital_gw1 <- GW1 |>
  filter(`hospital/non-hospital` == "Hospital") |>
  count(GW1) |>
  mutate(Hospital = (n / sum(n))*100)|>
  select(GW1, Hospital)

nonhospital_gw1 <- GW1 |>
  filter(`hospital/non-hospital` == "Non-Hospital") |>
  count(GW1) |>
  mutate(Nonhospital = (n / sum(n))*100)|>
  select(GW1, Nonhospital)

public_gw1 <- GW1 |>
  filter(`public/private` == "Public") |>
  count(GW1) |>
  mutate(Public = (n / sum(n))*100)|>
  select(GW1, Public)

private_gw1 <- GW1 |>
  filter(`public/private` == "Private") |>
  count(GW1) |>
  mutate(Private = (n / sum(n))*100)|>
  select(GW1, Private)

# joining and replace NA with zero 
water_source <- total_gw1 |>
  left_join(urban_gw1) |>
  left_join(rural_gw1) |>
  left_join(hospital_gw1) |>
  left_join(nonhospital_gw1) |>
  left_join(public_gw1) |>
  left_join(private_gw1)

# replacing NA with zero
water_source <- water_source |>
  replace(is.na(water_source), 0)

# calculate improved
improved_water_source <- water_source |>
  filter(GW1 == "Piped Supply Outside the building" |
         GW1 == "Piped Supply inside the building" |
         GW1 == "Protected dug well" |
         GW1 == "Tanker truck" |
         GW1 == "Rainwater"    |
         GW1 == "Tube well/Borehole"
        ) |>
  summarise_at(vars(Total, Urban, Rural, Hospital, Nonhospital, Public, Private), sum)|>
  mutate(Indicator = "Improved water source") |>
  relocate(Indicator, .before = Total)

# calculate no service
no_service_water <- water_source |>
  filter(GW1 == "Don't Know" |
         GW1 == "No water source or patients bring water from home" |
         GW1 == "Other" |
         GW1 == "Unprotected dug well"
        ) |>
  summarise_at(vars(Total, Urban, Rural, Hospital, Nonhospital, Public, Private), sum) |>
  mutate(Indicator = "No service") |>
  relocate(Indicator, .before = Total)

```

### On premises

Calculating on premises indicator for each stratifier

```{r}

# calculate on premises
GW2 <- data |>
  select(GW2, `urban/rural`, `hospital/non-hospital`, `public/private`)

total_gw2 <- GW2 |>
  count(GW2) |>
  mutate(Total = (n / sum(n))*100) |>
  select(GW2, Total) |>
  filter(GW2 == "On premises")

urban_gw2 <- GW2 |>
  filter(`urban/rural`== "Urban") |>
  count(GW2) |>
  mutate(Urban = (n / sum(n))*100)|>
  select(GW2, Urban) |>
  filter(GW2 == "On premises")

rural_gw2 <- GW2 |>
  filter(`urban/rural`== "Rural") |>
  count(GW2) |>
  mutate(Rural = (n / sum(n))*100)|>
  select(GW2, Rural)|>
  filter(GW2 == "On premises")

hospital_gw2 <- GW2 |>
  filter(`hospital/non-hospital` == "Hospital") |>
  count(GW2) |>
  mutate(Hospital = (n / sum(n))*100)|>
  select(GW2, Hospital)|>
  filter(GW2 == "On premises")

nonhospital_gw2 <- GW2 |>
  filter(`hospital/non-hospital` == "Non-Hospital") |>
  count(GW2) |>
  mutate(Nonhospital = (n / sum(n))*100)|>
  select(GW2, Nonhospital)|>
  filter(GW2 == "On premises")

public_gw2 <- GW2 |>
  filter(`public/private` == "Public") |>
  count(GW2) |>
  mutate(Public = (n / sum(n))*100)|>
  select(GW2, Public)|>
  filter(GW2 == "On premises")

private_gw2 <- GW2 |>
  filter(`public/private` == "Private") |>
  count(GW2) |>
  mutate(Private = (n / sum(n))*100)|>
  select(GW2, Private)|>
  filter(GW2 == "On premises")

# joining
on_premises <- total_gw2 |>
  left_join(urban_gw2) |>
  left_join(rural_gw2) |>
  left_join(hospital_gw2) |>
  left_join(nonhospital_gw2) |>
  left_join(public_gw2) |>
  left_join(private_gw2) |>
  rename(Indicator = GW2)

```

### Availability

Calculating water availability indicator for each stratifier

```{r}

# calculate availability
GW3 <- data |>
  select(GW3, `urban/rural`, `hospital/non-hospital`, `public/private`)

total_gw3 <- GW3 |>
  count(GW3) |>
  mutate(Total = (n / sum(n))*100) |>
  select(GW3, Total) |>
  filter(GW3 == "Yes")

urban_gw3 <- GW3 |>
  filter(`urban/rural`== "Urban") |>
  count(GW3) |>
  mutate(Urban = (n / sum(n))*100)|>
  select(GW3, Urban) |>
  filter(GW3 == "Yes")

rural_gw3 <- GW3 |>
  filter(`urban/rural`== "Rural") |>
  count(GW3) |>
  mutate(Rural = (n / sum(n))*100)|>
  select(GW3, Rural)|>
  filter(GW3 == "Yes")

hospital_gw3 <- GW3 |>
  filter(`hospital/non-hospital` == "Hospital") |>
  count(GW3) |>
  mutate(Hospital = (n / sum(n))*100)|>
  select(GW3, Hospital)|>
  filter(GW3 == "Yes")

nonhospital_gw3 <- GW3 |>
  filter(`hospital/non-hospital` == "Non-Hospital") |>
  count(GW3) |>
  mutate(Nonhospital = (n / sum(n))*100)|>
  select(GW3, Nonhospital)|>
  filter(GW3 == "Yes")

public_gw3 <- GW3 |>
  filter(`public/private` == "Public") |>
  count(GW3) |>
  mutate(Public = (n / sum(n))*100)|>
  select(GW3, Public)|>
  filter(GW3 == "Yes")

private_gw3 <- GW3 |>
  filter(`public/private` == "Private") |>
  count(GW3) |>
  mutate(Private = (n / sum(n))*100)|>
  select(GW3, Private)|>
  filter(GW3 == "Yes")

# joining
available <- total_gw3 |>
  left_join(urban_gw3) |>
  left_join(rural_gw3) |>
  left_join(hospital_gw3) |>
  left_join(nonhospital_gw3) |>
  left_join(public_gw3) |>
  left_join(private_gw3) |>
  mutate(Indicator = "Available",
         GW3= NULL) |>
  relocate(Indicator, .before = Total)

```

### Improved and available

Calculating improved and available water indicator for each stratifier

```{r}

# calculating improved and available
GW13 <- data |>
  select(GW1, GW3, `urban/rural`, `hospital/non-hospital`, `public/private`)

total_gw13 <- GW13 |>
  filter(GW3 == "Yes") |>
  filter(GW1 == "Piped Supply Outside the building" |
         GW1 == "Piped Supply inside the building" |
         GW1 == "Protected dug well" |
         GW1 == "Tanker truck" |
         GW1 == "Rainwater"    |
         GW1 == "Tube well/Borehole") |>
  count(GW3) |>
  mutate(Total = (n / total_sample)*100) |>
  select(GW3, Total)

urban_gw13 <- GW13 |>
  filter(`urban/rural`== "Urban") |>
  filter(GW3 == "Yes") |>
  filter(GW1 == "Piped Supply Outside the building" |
         GW1 == "Piped Supply inside the building" |
         GW1 == "Protected dug well" |
         GW1 == "Tanker truck" |
         GW1 == "Rainwater"    |
         GW1 == "Tube well/Borehole") |>
  count(GW3) |>
  mutate(Urban = (n / total_urban)*100) |>
  select(GW3, Urban)

rural_gw13 <- GW13 |>
  filter(`urban/rural`== "Rural") |>
  filter(GW3 == "Yes") |>
  filter(GW1 == "Piped Supply Outside the building" |
         GW1 == "Piped Supply inside the building" |
         GW1 == "Protected dug well" |
         GW1 == "Tanker truck" |
         GW1 == "Rainwater"    |
         GW1 == "Tube well/Borehole") |>
  count(GW3) |>
  mutate(Rural = (n / total_rural)*100) |>
  select(GW3, Rural)


hospital_gw13 <- GW13 |>
  filter(`hospital/non-hospital`== "Hospital") |>
  filter(GW3 == "Yes") |>
  filter(GW1 == "Piped Supply Outside the building" |
         GW1 == "Piped Supply inside the building" |
         GW1 == "Protected dug well" |
         GW1 == "Tanker truck" |
         GW1 == "Rainwater"    |
         GW1 == "Tube well/Borehole") |>
  count(GW3) |>
  mutate(Hospital = (n / total_hospital)*100) |>
  select(GW3, Hospital)

nonhospital_gw13 <- GW13 |>
  filter(`hospital/non-hospital`== "Non-Hospital") |>
  filter(GW3 == "Yes") |>
  filter(GW1 == "Piped Supply Outside the building" |
         GW1 == "Piped Supply inside the building" |
         GW1 == "Protected dug well" |
         GW1 == "Tanker truck" |
         GW1 == "Rainwater"    |
         GW1 == "Tube well/Borehole") |>
  count(GW3) |>
  mutate(Nonhospital = (n / total_nonhospital)*100) |>
  select(GW3, Nonhospital)

public_gw13 <- GW13 |>
  filter(`public/private`== "Public") |>
  filter(GW3 == "Yes") |>
  filter(GW1 == "Piped Supply Outside the building" |
         GW1 == "Piped Supply inside the building" |
         GW1 == "Protected dug well" |
         GW1 == "Tanker truck" |
         GW1 == "Rainwater"    |
         GW1 == "Tube well/Borehole") |>
  count(GW3) |>
  mutate(Public = (n / total_public)*100) |>
  select(GW3, Public)

private_gw13 <- GW13 |>
  filter(`public/private`== "Private") |>
  filter(GW3 == "Yes") |>
  filter(GW1 == "Piped Supply Outside the building" |
         GW1 == "Piped Supply inside the building" |
         GW1 == "Protected dug well" |
         GW1 == "Tanker truck" |
         GW1 == "Rainwater"    |
         GW1 == "Tube well/Borehole") |>
  count(GW3) |>
  mutate(Private = (n / total_private)*100) |>
  select(GW3, Private)

# joining
improved_available <- total_gw13 |>
  left_join(urban_gw13) |>
  left_join(rural_gw13) |>
  left_join(hospital_gw13) |>
  left_join(nonhospital_gw13) |>
  left_join(public_gw13) |>
  left_join(private_gw13) |>
  mutate(Indicator = "Improved and available",
         GW3 = NULL,
         Total =  Total$n,
         Urban = Urban$n,
         Rural = Rural$n,
         Hospital = Hospital$n,
         Nonhospital = Nonhospital$n,
         Public = Public$n,
         Private = Private$n) |>
  relocate(Indicator, .before = Total)

```

### Improved and on premises

Calculating improved and on premises indicator for each stratifier

```{r}

#calculating improved and on premises
GW12 <- data |>
  select(GW1, GW2, `urban/rural`, `hospital/non-hospital`, `public/private`)

total_gw12 <- GW12 |>
  filter(GW2 == "On premises") |>
  filter(GW1 == "Piped Supply Outside the building" |
         GW1 == "Piped Supply inside the building" |
         GW1 == "Protected dug well" |
         GW1 == "Tanker truck" |
         GW1 == "Rainwater"    |
         GW1 == "Tube well/Borehole") |>
  count(GW2) |>
  mutate(Total = (n / total_sample)*100) |>
  select(GW2, Total)

urban_gw12 <- GW12 |>
  filter(`urban/rural`== "Urban") |>
  filter(GW2 == "On premises") |>
  filter(GW1 == "Piped Supply Outside the building" |
         GW1 == "Piped Supply inside the building" |
         GW1 == "Protected dug well" |
         GW1 == "Tanker truck" |
         GW1 == "Rainwater"    |
         GW1 == "Tube well/Borehole") |>
  count(GW2) |>
  mutate(Urban = (n / total_urban)*100) |>
  select(GW2, Urban)

rural_gw12 <- GW12 |>
  filter(`urban/rural`== "Rural") |>
  filter(GW2 == "On premises") |>
  filter(GW1 == "Piped Supply Outside the building" |
         GW1 == "Piped Supply inside the building" |
         GW1 == "Protected dug well" |
         GW1 == "Tanker truck" |
         GW1 == "Rainwater"    |
         GW1 == "Tube well/Borehole") |>
  count(GW2) |>
  mutate(Rural = (n / total_rural)*100) |>
  select(GW2, Rural)

hospital_gw12 <- GW12 |>
  filter(`hospital/non-hospital`== "Hospital") |>
  filter(GW2 == "On premises") |>
  filter(GW1 == "Piped Supply Outside the building" |
         GW1 == "Piped Supply inside the building" |
         GW1 == "Protected dug well" |
         GW1 == "Tanker truck" |
         GW1 == "Rainwater"    |
         GW1 == "Tube well/Borehole") |>
  count(GW2) |>
  mutate(Hospital = (n / total_hospital)*100) |>
  select(GW2, Hospital)

nonhospital_gw12 <- GW12 |>
  filter(`hospital/non-hospital`== "Non-Hospital") |>
  filter(GW2 == "On premises") |>
  filter(GW1 == "Piped Supply Outside the building" |
         GW1 == "Piped Supply inside the building" |
         GW1 == "Protected dug well" |
         GW1 == "Tanker truck" |
         GW1 == "Rainwater"    |
         GW1 == "Tube well/Borehole") |>
  count(GW2) |>
  mutate(Nonhospital = (n / total_nonhospital)*100) |>
  select(GW2, Nonhospital)

public_gw12 <- GW12 |>
  filter(`public/private`== "Public") |>
  filter(GW2 == "On premises") |>
  filter(GW1 == "Piped Supply Outside the building" |
         GW1 == "Piped Supply inside the building" |
         GW1 == "Protected dug well" |
         GW1 == "Tanker truck" |
         GW1 == "Rainwater"    |
         GW1 == "Tube well/Borehole") |>
  count(GW2) |>
  mutate(Public = (n / total_public)*100) |>
  select(GW2, Public)

private_gw12 <- GW12 |>
  filter(`public/private`== "Private") |>
  filter(GW2 == "On premises") |>
  filter(GW1 == "Piped Supply Outside the building" |
         GW1 == "Piped Supply inside the building" |
         GW1 == "Protected dug well" |
         GW1 == "Tanker truck" |
         GW1 == "Rainwater"    |
         GW1 == "Tube well/Borehole") |>
  count(GW2) |>
  mutate(Private = (n / total_private)*100) |>
  select(GW2, Private)

# joining
improved_onpremises <- total_gw12 |>
  left_join(urban_gw12) |>
  left_join(rural_gw12) |>
  left_join(hospital_gw12) |>
  left_join(nonhospital_gw12) |>
  left_join(public_gw12) |>
  left_join(private_gw12) |>
  mutate(Indicator = "Improved and on premises",
         GW2 = NULL,
         Total =  Total$n,
         Urban = Urban$n,
         Rural = Rural$n,
         Hospital = Hospital$n,
         Nonhospital = Nonhospital$n,
         Public = Public$n,
         Private = Private$n) |>
  relocate(Indicator, .before = Total)

```

### Basic and limited services

Calculating basic and limited service for each stratifier

```{r}

#calculating basic
GW123 <- data |>
  select(GW1, GW2, GW3, `urban/rural`, `hospital/non-hospital`, `public/private`)

basic_total <- GW123 |>
  filter(GW2 == "On premises") |>
  filter(GW3 == "Yes") |>
  filter(GW1 == "Piped Supply Outside the building" |
         GW1 == "Piped Supply inside the building" |
         GW1 == "Protected dug well" |
         GW1 == "Tanker truck" |
         GW1 == "Rainwater"    |
         GW1 == "Tube well/Borehole") |>
  count(GW2) |>
  mutate(Total = (n / total_sample)*100) |>
  select(GW2, Total)

basic_urban <- GW123 |>
  filter(`urban/rural`== "Urban") |>
  filter(GW2 == "On premises") |>
  filter(GW3 == "Yes") |>
  filter(GW1 == "Piped Supply Outside the building" |
         GW1 == "Piped Supply inside the building" |
         GW1 == "Protected dug well" |
         GW1 == "Tanker truck" |
         GW1 == "Rainwater"    |
         GW1 == "Tube well/Borehole") |>
  count(GW2) |>
  mutate(Urban = (n / total_urban)*100) |>
  select(GW2, Urban)

basic_rural <- GW123 |>
  filter(`urban/rural`== "Rural") |>
  filter(GW2 == "On premises") |>
  filter(GW3 == "Yes") |>
  filter(GW1 == "Piped Supply Outside the building" |
         GW1 == "Piped Supply inside the building" |
         GW1 == "Protected dug well" |
         GW1 == "Tanker truck" |
         GW1 == "Rainwater"    |
         GW1 == "Tube well/Borehole") |>
  count(GW2) |>
  mutate(Rural = (n / total_rural)*100) |>
  select(GW2, Rural)

basic_hospital <- GW123 |>
  filter(`hospital/non-hospital`== "Hospital") |>
  filter(GW2 == "On premises") |>
  filter(GW3 == "Yes") |>
  filter(GW1 == "Piped Supply Outside the building" |
         GW1 == "Piped Supply inside the building" |
         GW1 == "Protected dug well" |
         GW1 == "Tanker truck" |
         GW1 == "Rainwater"    |
         GW1 == "Tube well/Borehole") |>
  count(GW2) |>
  mutate(Hospital = (n / total_hospital)*100) |>
  select(GW2, Hospital)

basic_nonhospital <- GW123 |>
  filter(`hospital/non-hospital`== "Non-Hospital") |>
  filter(GW2 == "On premises") |>
  filter(GW3 == "Yes") |>
  filter(GW1 == "Piped Supply Outside the building" |
         GW1 == "Piped Supply inside the building" |
         GW1 == "Protected dug well" |
         GW1 == "Tanker truck" |
         GW1 == "Rainwater"    |
         GW1 == "Tube well/Borehole") |>
  count(GW2) |>
  mutate(Nonhospital = (n / total_nonhospital)*100) |>
  select(GW2, Nonhospital)

basic_public <- GW123 |>
  filter(`public/private`== "Public") |>
  filter(GW2 == "On premises") |>
  filter(GW3 == "Yes") |>
  filter(GW1 == "Piped Supply Outside the building" |
         GW1 == "Piped Supply inside the building" |
         GW1 == "Protected dug well" |
         GW1 == "Tanker truck" |
         GW1 == "Rainwater"    |
         GW1 == "Tube well/Borehole") |>
  count(GW2) |>
  mutate(Public = (n / total_public)*100) |>
  select(GW2, Public)

basic_private <- GW123 |>
  filter(`public/private`== "Private") |>
  filter(GW2 == "On premises") |>
  filter(GW3 == "Yes") |>
  filter(GW1 == "Piped Supply Outside the building" |
         GW1 == "Piped Supply inside the building" |
         GW1 == "Protected dug well" |
         GW1 == "Tanker truck" |
         GW1 == "Rainwater"    |
         GW1 == "Tube well/Borehole") |>
  count(GW2) |>
  mutate(Private = (n / total_private)*100) |>
  select(GW2, Private)

# joining
basic_water <- basic_total |>
  left_join(basic_urban) |>
  left_join(basic_rural) |>
  left_join(basic_hospital) |>
  left_join(basic_nonhospital) |>
  left_join(basic_public) |>
  left_join(basic_private) |>
  mutate(Indicator = "Basic",
         GW2 = NULL,
         Total =  Total$n,
         Urban = Urban$n,
         Rural = Rural$n,
         Hospital = Hospital$n,
         Nonhospital = Nonhospital$n,
         Public = Public$n,
         Private = Private$n) |>
  relocate(Indicator, .before = Total)

# calculating limited
limited_water <- tibble(
  Total = 100 - (basic_water[2] + no_service_water[2]),
  Urban = 100 - (basic_water[3] + no_service_water[3]),
  Rural = 100 - (basic_water[4] + no_service_water[4]),
  Hospital = 100 - (basic_water[5] + no_service_water[5]),
  Nonhospital = 100 - (basic_water[6] + no_service_water[6]),
  Public = 100 - (basic_water[7] + no_service_water[7]),
  Private = 100 - (basic_water[8] + no_service_water[8])) |>
  mutate(Indicator = "Limited",
         Total =  Total$Total,
         Urban = Urban$Urban,
         Rural = Rural$Rural,
         Hospital = Hospital$Hospital,
         Nonhospital = Nonhospital$Nonhospital,
         Public = Public$Public,
         Private = Private$Private) |>
  relocate(Indicator, .before = Total)

```

## Hygiene Data

### Points of care

Calculating handwashing facilities with water and soap at points of care for each stratifier

```{r}

# calculating handwashing facilities with water and soap at points of care
GH1 <- data |>
  select(GH1, `urban/rural`, `hospital/non-hospital`, `public/private`) |>
  filter(GH1 == "Yes")

total_gh1 <- GH1 |>
  count(GH1) |>
  mutate(Total = (n / total_sample)*100) |>
  select(GH1, Total)

urban_gh1 <- GH1 |>
  filter(`urban/rural`== "Urban") |>
  count(GH1) |>
  mutate(Urban = (n / total_urban)*100)|>
  select(GH1, Urban)

rural_gh1 <- GH1 |>
  filter(`urban/rural`== "Rural") |>
  count(GH1) |>
  mutate(Rural = (n / total_rural)*100)|>
  select(GH1, Rural)

hospital_gh1 <- GH1 |>
  filter(`hospital/non-hospital` == "Hospital") |>
  count(GH1) |>
  mutate(Hospital = (n / total_hospital)*100)|>
  select(GH1, Hospital)

nonhospital_gh1 <- GH1 |>
  filter(`hospital/non-hospital` == "Non-Hospital") |>
  count(GH1) |>
  mutate(Nonhospital = (n / total_nonhospital)*100)|>
  select(GH1, Nonhospital)

public_gh1 <- GH1 |>
  filter(`public/private` == "Public") |>
  count(GH1) |>
  mutate(Public = (n / total_public)*100)|>
  select(GH1, Public)

private_gh1 <- GH1 |>
  filter(`public/private` == "Private") |>
  count(GH1) |>
  mutate(Private = (n / total_private)*100)|>
  select(GH1, Private)

# joining and replace NA with zero 
pointsofcare <- total_gh1 |>
  left_join(urban_gh1) |>
  left_join(rural_gh1) |>
  left_join(hospital_gh1) |>
  left_join(nonhospital_gh1) |>
  left_join(public_gh1) |>
  left_join(private_gh1) |>
  mutate(Indicator = "Handwashing facilities with water and soap at points of care",
         GH1 = NULL,
         Total =  Total$n,
         Urban = Urban$n,
         Rural = Rural$n,
         Hospital = Hospital$n,
         Nonhospital = Nonhospital$n,
         Public = Public$n,
         Private = Private$n) |>
  relocate(Indicator, .before = Total)

```

### Toilets

Calculating handwashing facilities within 5 meters of toilets for each stratifier

```{r}

# calculating handwashing facilities within 5 metres of toilets
GH2 <- data |>
  select(GH2, `urban/rural`, `hospital/non-hospital`, `public/private`)

total_gh2 <- GH2 |>
  filter(GH2 == "Yes" |
         GH2 == "No, there are handwashing facilities near the toilets but lacking soap and /or water") |>
  count(GH2) |>
  summarise_at(vars(n), sum) |>
  mutate(GH2 = "Yes",
         Total = (n / total_sample) * 100,
         n = NULL)

urban_gh2 <- GH2 |>
  filter(`urban/rural`== "Urban",
         GH2 == "Yes" |
         GH2 == "No, there are handwashing facilities near the toilets but lacking soap and /or water") |>
  count(GH2) |>
  summarise_at(vars(n), sum) |>
  mutate(GH2 = "Yes",
         Urban = (n / total_urban) * 100,
         n = NULL)

rural_gh2 <- GH2 |>
  filter(`urban/rural`== "Rural",
         GH2 == "Yes" |
         GH2 == "No, there are handwashing facilities near the toilets but lacking soap and /or water") |>
  count(GH2) |>
  summarise_at(vars(n), sum) |>
  mutate(GH2 = "Yes",
         Rural = (n / total_rural) * 100,
         n = NULL)

hospital_gh2 <- GH2 |>
  filter(`hospital/non-hospital` == "Hospital",
         GH2 == "Yes" |
         GH2 == "No, there are handwashing facilities near the toilets but lacking soap and /or water") |>
  count(GH2) |>
  summarise_at(vars(n), sum) |>
  mutate(GH2 = "Yes",
         Hospital = (n / total_hospital) * 100,
         n = NULL)

nonhospital_gh2 <- GH2 |>
  filter(`hospital/non-hospital` == "Non-Hospital",
         GH2 == "Yes" |
         GH2 == "No, there are handwashing facilities near the toilets but lacking soap and /or water") |>
  count(GH2) |>
  summarise_at(vars(n), sum) |>
  mutate(GH2 = "Yes",
         Nonhospital = (n / total_nonhospital) * 100,
         n = NULL)

public_gh2 <- GH2 |>
  filter(`public/private` == "Public",
         GH2 == "Yes" |
         GH2 == "No, there are handwashing facilities near the toilets but lacking soap and /or water") |>
  count(GH2) |>
  summarise_at(vars(n), sum) |>
  mutate(GH2 = "Yes",
         Public = (n / total_public) * 100,
         n = NULL)

private_gh2 <- GH2 |>
  filter(`public/private` == "Private",
         GH2 == "Yes" |
         GH2 == "No, there are handwashing facilities near the toilets but lacking soap and /or water") |>
  count(GH2) |>
  summarise_at(vars(n), sum) |>
  mutate(GH2 = "Yes",
         Private = (n / total_private) * 100,
         n = NULL)

# joining
toilets <- total_gh2 |>
  left_join(urban_gh2) |>
  left_join(rural_gh2) |>
  left_join(hospital_gh2) |>
  left_join(nonhospital_gh2) |>
  left_join(public_gh2) |>
  left_join(private_gh2) |>
  mutate(Indicator = "Handwashing facilities within 5 metres of toilets",
         GH2 = NULL,
         Total =  Total$n,
         Urban = Urban$n,
         Rural = Rural$n,
         Hospital = Hospital$n,
         Nonhospital = Nonhospital$n,
         Public = Public$n,
         Private = Private$n) |>
  relocate(Indicator, .before = Total)

# calculating handwashing facilities within 5 metres of toilets with water and soap
GH2_fun <- data |>
  select(GH2, `urban/rural`, `hospital/non-hospital`, `public/private`) |>
  filter(GH2 == "Yes")

total_gh2_fun <- GH2_fun |>
  count(GH2) |>
  mutate(Total = (n / total_sample)*100) |>
  select(GH2, Total)

urban_gh2_fun <- GH2_fun |>
  filter(`urban/rural`== "Urban") |>
  count(GH2) |>
  mutate(Urban = (n / total_urban)*100)|>
  select(GH2, Urban)

rural_gh2_fun <- GH2_fun |>
  filter(`urban/rural`== "Rural") |>
  count(GH2) |>
  mutate(Rural = (n / total_rural)*100)|>
  select(GH2, Rural)

hospital_gh2_fun <- GH2_fun |>
  filter(`hospital/non-hospital` == "Hospital") |>
  count(GH2) |>
  mutate(Hospital = (n / total_hospital)*100)|>
  select(GH2, Hospital)

nonhospital_gh2_fun <- GH2_fun |>
  filter(`hospital/non-hospital` == "Non-Hospital") |>
  count(GH2) |>
  mutate(Nonhospital = (n / total_nonhospital)*100)|>
  select(GH2, Nonhospital)

public_gh2_fun <- GH2_fun |>
  filter(`public/private` == "Public") |>
  count(GH2) |>
  mutate(Public = (n / total_public)*100)|>
  select(GH2, Public)

private_gh2_fun <- GH2_fun |>
  filter(`public/private` == "Private") |>
  count(GH2) |>
  mutate(Private = (n / total_private)*100)|>
  select(GH2, Private)

# joining
toilets_fun <- total_gh2_fun |>
  left_join(urban_gh2_fun) |>
  left_join(rural_gh2_fun) |>
  left_join(hospital_gh2_fun) |>
  left_join(nonhospital_gh2_fun) |>
  left_join(public_gh2_fun) |>
  left_join(private_gh2_fun) |>
  mutate(Indicator = "Handwashing facilities within 5 metres of toilets with water and soap",
         GH2 = NULL,
         Total =  Total$n,
         Urban = Urban$n,
         Rural = Rural$n,
         Hospital = Hospital$n,
         Nonhospital = Nonhospital$n,
         Public = Public$n,
         Private = Private$n) |>
  relocate(Indicator, .before = Total)

```

### Basic and limited services

Calculating basic and limited service for each stratifier

```{r}

# calculating basic
GH12 <- data |>
  select(GH1, GH2, `urban/rural`, `hospital/non-hospital`, `public/private`) |>
  filter(GH1 == "Yes" &
         GH2 == "Yes")

total_gh12 <- GH12 |>
  count(GH2) |>
  mutate(Total = (n / total_sample)*100) |>
  select(GH2, Total)

urban_gh12 <- GH12 |>
  filter(`urban/rural`== "Urban") |>
  count(GH2) |>
  mutate(Urban = (n / total_urban)*100)|>
  select(GH2, Urban)

rural_gh12 <- GH12 |>
  filter(`urban/rural`== "Rural") |>
  count(GH2) |>
  mutate(Rural = (n / total_rural)*100)|>
  select(GH2, Rural)

hospital_gh12 <- GH12 |>
  filter(`hospital/non-hospital` == "Hospital") |>
  count(GH2) |>
  mutate(Hospital = (n / total_hospital)*100)|>
  select(GH2, Hospital)

nonhospital_gh12 <- GH12 |>
  filter(`hospital/non-hospital` == "Non-Hospital") |>
  count(GH2) |>
  mutate(Nonhospital = (n / total_nonhospital)*100)|>
  select(GH2, Nonhospital)

public_gh12 <- GH12 |>
  filter(`public/private` == "Public") |>
  count(GH2) |>
  mutate(Public = (n / total_public)*100)|>
  select(GH2, Public)

private_gh12 <- GH12 |>
  filter(`public/private` == "Private") |>
  count(GH2) |>
  mutate(Private = (n / total_private)*100)|>
  select(GH2, Private)

# joining
basic_hygiene <- total_gh12 |>
  left_join(urban_gh12) |>
  left_join(rural_gh12) |>
  left_join(hospital_gh12) |>
  left_join(nonhospital_gh12) |>
  left_join(public_gh12) |>
  left_join(private_gh12) |>
  mutate(Indicator = "Basic",
         GH2 = NULL,
         Total =  Total$n,
         Urban = Urban$n,
         Rural = Rural$n,
         Hospital = Hospital$n,
         Nonhospital = Nonhospital$n,
         Public = Public$n,
         Private = Private$n) |>
  relocate(Indicator, .before = Total)

# calculating no serivce
GH12_no <- data |>
  select(GH1, GH2, `urban/rural`, `hospital/non-hospital`, `public/private`)

total_gh12_no <- GH12_no |>
  filter(GH1 != "Yes" &
         GH2 != "Yes") |>
  count(GH2) |>
  summarise_at(vars(n), sum) |>
  mutate(GH2 = "Yes",
         Total = (n / total_sample) * 100,
         n = NULL)

urban_gh12_no <- GH12_no |>
  filter(`urban/rural`== "Urban") |>
  filter(GH1 != "Yes" &
         GH2 != "Yes") |>
  count(GH2) |>
  summarise_at(vars(n), sum) |>
  mutate(GH2 = "Yes",
         Urban = (n / total_urban) * 100,
         n = NULL)

rural_gh12_no <- GH12_no |>
  filter(`urban/rural`== "Rural") |>
  filter(GH1 != "Yes" &
         GH2 != "Yes") |>
  count(GH2) |>
  summarise_at(vars(n), sum) |>
  mutate(GH2 = "Yes",
         Rural = (n / total_rural) * 100,
         n = NULL)

hospital_gh12_no <- GH12_no |>
  filter(`hospital/non-hospital` == "Hospital") |>
  filter(GH1 != "Yes" &
         GH2 != "Yes") |>
  count(GH2) |>
  summarise_at(vars(n), sum) |>
  mutate(GH2 = "Yes",
         Hospital = (n / total_hospital) * 100,
         n = NULL)

nonhospital_gh12_no <- GH12_no |>
  filter(`hospital/non-hospital` == "Non-Hospital") |>
  filter(GH1 != "Yes" &
         GH2 != "Yes") |>
  count(GH2) |>
  summarise_at(vars(n), sum) |>
  mutate(GH2 = "Yes",
         Nonhospital = (n / total_nonhospital) * 100,
         n = NULL)

public_gh12_no <- GH12_no |>
  filter(`public/private` == "Public") |>
  filter(GH1 != "Yes" &
         GH2 != "Yes") |>
  count(GH2) |>
  summarise_at(vars(n), sum) |>
  mutate(GH2 = "Yes",
         Public = (n / total_public) * 100,
         n = NULL)

private_gh12_no <- GH12_no |>
  filter(`public/private` == "Private") |>
  filter(GH1 != "Yes" &
         GH2 != "Yes") |>
  count(GH2) |>
  summarise_at(vars(n), sum) |>
  mutate(GH2 = "Yes",
         Private = (n / total_private) * 100,
         n = NULL)

# joining
no_service_hygiene <- total_gh12_no |>
  left_join(urban_gh12_no) |>
  left_join(rural_gh12_no) |>
  left_join(hospital_gh12_no) |>
  left_join(nonhospital_gh12_no) |>
  left_join(public_gh12_no) |>
  left_join(private_gh12_no) |>
  mutate(Indicator = "No service",
         GH2 = NULL,
         Total =  Total$n,
         Urban = Urban$n,
         Rural = Rural$n,
         Hospital = Hospital$n,
         Nonhospital = Nonhospital$n,
         Public = Public$n,
         Private = Private$n) |>
  relocate(Indicator, .before = Total)

# calculating limited
limited_hygiene <- tibble(
  Total = 100 - (basic_hygiene[2] + no_service_hygiene[2]),
  Urban = 100 - (basic_hygiene[3] + no_service_hygiene[3]),
  Rural = 100 - (basic_hygiene[4] + no_service_hygiene[4]),
  Hospital = 100 - (basic_hygiene[5] + no_service_hygiene[5]),
  Nonhospital = 100 - (basic_hygiene[6] + no_service_hygiene[6]),
  Public = 100 - (basic_hygiene[7] + no_service_hygiene[7]),
  Private = 100 - (basic_hygiene[8] + no_service_hygiene[8])) |>
  mutate(Indicator = "Limited",
         Total =  Total$Total,
         Urban = Urban$Urban,
         Rural = Rural$Rural,
         Hospital = Hospital$Hospital,
         Nonhospital = Nonhospital$Nonhospital,
         Public = Public$Public,
         Private = Private$Private) |>
  relocate(Indicator, .before = Total)

```

# Results

## Water indicators

See @tbl-water summarizes all water indicators resulted from the analysis.

Significant disparities in water indicators exist in terms of the types of facilities; facilities located in urban areas as well as hospitals are more likely to meet the criteria for basic water service than facilities located in rural areas and smaller health care facilities, respectively.

```{r}
#| label: tbl-water
#| tbl-cap: Water indicators

improved_water_source |>
  full_join(on_premises) |>
  full_join(available) |>
  full_join(improved_available) |>
  full_join(improved_onpremises) |>
  full_join(basic_water) |>
  full_join(limited_water) |>
  full_join(no_service_water) |>
  kable(digits = 1)

```

## Hygiene indicators

See @tbl-hygiene summarizes all hygiene indicators resulted from the analysis.

Significant disparities in hygiene indicators exist in terms of the types of facilities; facilities located in urban areas as well as hospitals are more likely to meet the criteria for basic hygiene service than facilities located in rural areas and smaller health care facilities, respectively.

```{r}
#| label: tbl-hygiene
#| tbl-cap: Hygiene indicators

pointsofcare |>
  full_join(toilets) |>
  full_join(toilets_fun) |>
  full_join(basic_hygiene) |>
  full_join(limited_hygiene) |>
  full_join(no_service_hygiene) |>
  kable(digits = 1)

```

# Conclusion

## Basic water ladder

See @fig-water illustrates basic water ladder resulted from the analysis. Significant disparities are depicted in the basic water ladders across all settings.

```{r}
#| label: fig-water
#| fig-cap: Basic water ladder

# create object
basic_water_ladder <-
  (basic_water) |>
  full_join(limited_water) |>
  full_join(no_service_water)

# pivot
basic_water_ladder <-
  basic_water_ladder |>
  pivot_longer(cols = Total:Private,
             names_to = "Setting",
             values_to = "Percent")

# reorder indicators
basic_water_ladder$Indicator <- factor(basic_water_ladder$Indicator, levels = c("No service", "Limited", "Basic"))

# reorder settings
basic_water_ladder$Setting <- factor(basic_water_ladder$Setting, levels = c("Total", "Urban", "Rural", "Hospital", "Nonhospital", "Public", "Private"))

# Assigning white for basic in geom_text
basic_water_ladder <- basic_water_ladder |>
  mutate(Color = case_when(
    Indicator == "Basic" ~ "white",
    .default = "black"))

# plot water ladders
ggplot(data = basic_water_ladder,
       mapping = aes(x = Setting,
                     y = Percent,
                     fill = Indicator)) +
  geom_col() +
  geom_text(aes(label = round(Percent, 0)),
            position = position_stack(vjust = 0.5),
            size = 3,
            color = basic_water_ladder$Color) +
  theme_minimal() +
  labs(title = "Water Indicator",
       subtitle = "JMP basic water service ladder",
       x = "Setting",
       y = "Proportion of healthcare facilities (%)") +
  scale_fill_manual(
    values = c("No service" = "#FEBC11", "Limited" = "#FFF176", "Basic" = "#00B8EC"),
    breaks = c("No service", "Limited","Basic")) +
  guides(fill = guide_legend(title = NULL)) +
  theme(panel.grid.minor.y = element_blank(),
        panel.grid.major.x = element_blank())

```

## Basic hygiene ladder

See @fig-hygiene illustrates basic hygiene ladder resulted from the analysis. Significant disparities are depicted in the basic hygiene ladders across all settings.

```{r}
#| label: fig-hygiene
#| fig-cap: Basic hygiene ladder

# create object
basic_hygiene_ladder <-
  (basic_hygiene) |>
  full_join(limited_hygiene) |>
  full_join(no_service_hygiene)

# pivot
basic_hygiene_ladder <-
  basic_hygiene_ladder |>
  pivot_longer(cols = Total:Private,
             names_to = "Setting",
             values_to = "Percent")

# reorder indicators
basic_hygiene_ladder$Indicator <- factor(basic_hygiene_ladder$Indicator, levels = c("No service", "Limited", "Basic"))

# reorder settings
basic_hygiene_ladder$Setting <- factor(basic_hygiene_ladder$Setting, levels = c("Total", "Urban", "Rural", "Hospital", "Nonhospital", "Public", "Private"))

# Assigning white for basic in geom_text
basic_hygiene_ladder <- basic_hygiene_ladder |>
  mutate(Color = case_when(
    Indicator == "Basic" ~ "white",
    .default = "black"))

# plot hygiene ladders
ggplot(data = basic_hygiene_ladder,
       mapping = aes(x = Setting,
                     y = Percent,
                     fill = Indicator)) +
  geom_col() +
  geom_text(aes(label = round(Percent, 0)),
            position = position_stack(vjust = 0.5),
            size = 3,
            color = basic_hygiene_ladder$Color) +
  theme_minimal() +
  labs(title = "Hygiene Indicator",
       subtitle = "JMP basic hygiene service ladder",
       x = "Setting",
       y = "Proportion of healthcare facilities (%)") +
  scale_fill_manual(
    values = c("No service" = "#FEBC11", "Limited" = "#FFF176", "Basic" = "#AB47BC"),
    breaks = c("No service", "Limited","Basic")) +
  guides(fill = guide_legend(title = NULL)) +
  theme(panel.grid.minor.y = element_blank(),
        panel.grid.major.x = element_blank())
```

# References

1.  @WHOUNICEF2017
2.  @WHOUNICEF2019
